---
title: "How to Subset Large Weather Data Files"
author: "Author: Kristina Riemer"
output: github_document
urlcolor: blue
---

## Data file information

This is for how to subset the NARR `all.nc` file. It is on Globus in the UA Field Scanner collection [here](https://app.globus.org/file-manager?origin_id=cbae3a96-d081-4951-bd1e-a8fda974cefa&origin_path=%2Fmet%2Fnarr%2Fthreehourly%2F). The file is ~650GB. This is not global, it is restricted to the northern hemisphere as below: 

 * Latitudinal range: 10.125 - 83.875 
 * Longitudinal range: -169.875 to -50.125

The entire global file is the CRUNCEP `all.nc`, which is ~1.3TB and in the same Globus collection [here](https://app.globus.org/file-manager?origin_id=cbae3a96-d081-4951-bd1e-a8fda974cefa&origin_path=%2Fmet%2Fcruncep%2F). We will need more data space on the HPC to subset this because the upper limit is currently 1TB. 

## Subsetting process

### 1. Copy entire data file onto HPC

First increase data space by requesting an xdisk allocation on the HPC. Go to [the HPC web interface](https://ood.hpc.arizona.edu/pun/sys/dashboard), and open up the terminal by clicking on Files -> Home Directory -> Open in terminal -> Ocelote. Use `xdisk -c create -m 1000` to get 1TB of space.  

Open up the `all.nc` file on Globus using the link above. Then open up the HPC collection, which is called `arizona#sdmz-dtn`. Navigate to the new terrabyte of space in this collection at `/xdisk/username/`. Use the "Start" button to initiate transfer, which will take about an hour. Also transfer `champaign.nc` if desired. 

### 2. Determine desired dimensions

Use the following plotting functionality in R to figure out which latitudes and longitudes to clip the weather data to. 

```{r message=FALSE, warning=FALSE}
library(ncdf4)
library(dplyr)
library(maps)
library(ggplot2)
```

The first example shows the four corners of a long latitudinal slice that includes area in both China and Russia. Change out the `lat` and `lon` values in the data frame and the filtered region in the `background_map` as desired. 

```{r}
chiruss_corners = expand.grid(lat = c(30.1, 70.1),
                              lon = c(87.1, 87.6))
background_map <- map_data("world") %>%
  filter(region == "China" | region == "Russia")
ggplot()  +
  geom_polygon(data = background_map, aes(x = long, y = lat, group = group), fill = NA, color = "black") +
  geom_point(data = chiruss_corners, aes(x = lon, y = lat), color = "red") +
  geom_path(color = "white") +
  coord_equal() +
  theme(legend.position = "none")
```

This plot shows the limits of the NARR `all.nc` file. 

```{r}
narr_all_corners = expand.grid(lat = c(10.125, 83.875), 
                             lon = c(-169.875, -50.125))
background_map <- map_data("world")
ggplot() +
  geom_polygon(data = background_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  geom_point(data = narr_all_corners, aes(x = lon, y = lat), color = "red")
```

### 3. Subset file on HPC

In local command line, log onto UA HPC by executing `ssh username@hpc.arizona.edu`, then entering password and two factor authentication info. Choose the ocelote machine. 

Now on login node, where no analysis should occur. Code can be run either using a PBS script or in an interactive session. We'll use the latter for subsetting, as shown below. 

Here we are subsetting the North American NARR all.nc into a wide longitudinal range across the United States that includes St. Louis. 

1. Start interactive session with an hour of time

```{shell}
qsub -I -N stl_subset -W group_list=dlebauer -q standard -l select=1:ncpus=1:mem=6gb -l walltime=1:0:0
```

2. Navigate to David's xdisk space where the original file is stored

```{shell}
cd /xdisk/dlebauer/mig2020/xdisk/kristinariemer/narr/
```

3. Do subsetting steps using NCO tools, for both space and time

```{shell}
ncks -O -d longitude,-110.9,-80.9 -d latitude,37.5,39.5 all.nc stl_space.nc
ncea -d time,112494.4,113589.9 stl_space.nc stl_space_time.nc
```

4. Once done, check for final file and end session

```{shell}
exit
```

The relevant NCO commands for other weather subsets are shown here: 

Champaign: 
```{shell}
ncks -O -d longitude,-88.9,-87.5 -d latitude,39.8,40.5 all.nc champaign.nc
```

Eastern Illinois: 
```{shell}
ncks -O -d longitude,-88.9,-87.5 -d latitude,38.8,41.5 all.nc eaill.nc
```

USA/Canada: 
```{shell}
ncks -O -d longitude,-108.5,-105.5 -d latitude,45.5,55.5 all.nc usca.nc
```

Russia/China (*this will only work with the CRUNCEP all.nc*):

```{shell}
ncks -O -d longitude,87.1,87.9 -d latitude,30.1,70.1 all.nc chiruss.nc
```

### 4. Copy to local machine using Globus

Once the desired subset has been generated, you will need to transfer the file from the HPC collection `arizona#sdmz-dtn` on Globus to your own endpoint. This will require downloading Globus Connect Personal and following [these instructions](https://www.globus.org/globus-connect-personal) to create an endpoint on your machine. 

Once this is ready, select the file on the HPC collection and click the "Start" button to transfer to your local selected location. This can take an hour for for files 3-4GB in size. 

### 5. Check that dimensions are correct

Use the following example R code to plot all the locations in a transferred data subset. Each location gets a randomly chosen value, not from the weather data itself. The `background_map` may need to be modified depending on where on the globe the values are. 

Champaign: 
```{r}
champaign <- nc_open("champaign.nc")
champaign_latlon <- expand.grid(lat = ncvar_get(champaign, "latitude"), 
                                lon = ncvar_get(champaign, "longitude"))
champaign_latlon$value <- sample(1:nrow(champaign_latlon), nrow(champaign_latlon))
background_map <- map_data("state") %>% 
  filter(region == "illinois")

ggplot() +
  geom_polygon(data = background_map, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = champaign_latlon, aes(x = lon, y = lat, fill = value)) +
  coord_quickmap()
```

Eastern Illinois: 
```{r}
eaill <- nc_open("eaill.nc")
eaill_latlon <- expand.grid(lat = ncvar_get(eaill, "latitude"), 
                                lon = ncvar_get(eaill, "longitude"))
eaill_latlon$value <- sample(1:nrow(eaill_latlon), nrow(eaill_latlon))
background_map <- map_data("state") %>% 
  filter(region == "illinois")

ggplot() +
  geom_polygon(data = background_map, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = eaill_latlon, aes(x = lon, y = lat, fill = value)) +
  coord_quickmap()
```

USA/Canada: 
```{r}
usca <- nc_open("usca.nc")
usca_latlon <- expand.grid(lat = ncvar_get(usca, "latitude"), 
                                lon = ncvar_get(usca, "longitude"))
usca_latlon$value <- sample(1:nrow(usca_latlon), nrow(usca_latlon))
background_map <- map_data("world") %>% 
  filter(region == "USA" | region == "Canada")

ggplot() +
  geom_polygon(data = background_map, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = usca_latlon, aes(x = lon, y = lat, fill = value)) +
  coord_quickmap() +
  lims(x = c(-150, -50), y = c(20, 80))
```
