---
title: "Simultaneous fitting of CO2 and light response curves"
author: "Jessica Guo"
output: github_document
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

Necessary functions. 
```{r}
source("/home/jessicaguo/pecan/modules/photosynthesis/R/fitA.R")
library(postjags)
library(dplyr)
library(ggplot2)
source("R/multiplot.R")
load("outputs/Bayesian_parameter_estimation/out_mcmc.Rdata")
```

The above function is modified from the 'PEcAn.photosynthesis' package to include a Collatz model for C4 photosynthesis and an option for data from the Licor 6800. Retained from the original model include options for fixed and random effects. The function has 5 arguments:
- flux.data, required, a dataframe of gas exchange data, labeled with a column to associate sampling units. Default is 'fname', but can be indicated in the model argument below. This column is used to identify random effects. 
- cov.data, default is NULL, a dataframe of covariates should the fixed effects be used. The column names are used in model formulas and should be succint. 
- model, default is NULL, a list of at least six components:
    - a.fixed, a.random: fixed and random effects for parameter alpha, quantum yield
    - V.fixed, V.random: fixed and random effects for parameter Vmax, maximum Rubisco capacity
    - k.fixed, k.random: fixed and random effects for parameter k, initial slope of CO2 response
    - match: the variable used identify sampling units for random effects and to match the gas-exchange and covariate data for fixed effects. 
    - n.iter: the number of MCMC interations (n.iter). Default is 5000
- pathway, either "C3" or "C4"
- licor, either "6400" or "6800"

Here, our goal is to test performance of fitA with the Collatz model using the Danforth data. Three sets of data inputs will be used and resulting parameters compared: ACi curve only,  AQ curve only, and ACi & AQ curves simultaneously. As an added check, these parameters will be compared to outputs from "parameter_estimation.Rmd"
Only the default model will be used here. 

First, extract, label, and combine the ACi and AQ data for each experiment with both datasets. 
```{r}
expDF <- read.csv("cleaned_data/experiments.csv",
                  colClasses = c("character", "numeric", rep("character",2), rep("numeric",2)))

ids <- unique(expDF$ID[which(expDF$A_Ci == 1 & expDF$A_Qin == 1)])

datlist <- list()
for(i in 1:length(ids)){
  aci <- read.csv(paste0("cleaned_data/ACi/A_Ci_curves_", ids[i], ".csv"), header=T)
  aci$type <- rep("ACi", nrow(aci))
  aci <- subset(aci, select = c("type", "rep", "A", "Pci", "Qin"))
  
  aq <- read.csv(paste0("cleaned_data/AQ/AQin_curves_", ids[i], ".csv"), header=T)
  aq$type <- rep("AQ", nrow(aq))
  aq <- subset(aq, select = c("type", "rep", "A", "Pci", "Qin"))
  
  datlist[[i]] <- rbind(aci, aq)
}
```

Each plant replicate will be modeled separately, rather than as a random effect. There are currently 10 datasets with both ACi and AQ data, each measured on 3 plant replicates. A total of 3 sets of thirty outputs wil be recorded, distinguishing 
```{r}
# Data types to select
types <- list("ACi", "AQ", c("ACi", "AQ"))

# Simple model, no fixed or random effects
model <- list(V.fixed = NULL, V.random = NULL, 
                  a.fixed = NULL, a.random = NULL, 
                  k.fixed = NULL, k.random = NULL, 
                  n.iter = 5000, match = "rep")
out <- list()
for(i in 1:length(ids)){ # 10 experiments
  out[[i]] <- list()
  for(j in 1:3){ # 3 sets of inputs: ACi, AQ, or both
    out[[i]][[j]] <- list()
    for(k in 1:3){ # 3 replicate plants per experiment
        flux.data <- subset(datlist[[i]], rep == paste0("plant_", k) & type %in% types[[j]])
        out[[i]][[j]][[k]] <- fitA(flux.data = flux.data, pathway = "C4", licor = "6800",
                                   model = model)
        print(paste0("Run ", i, "_", j, "_", k, " completed"))
    }
  }
}

# Save output list as .Rdata object
if(dir.exists("outputs/Bayesian_parameter_estimation") == F){
  dir.create("outputs/Bayesian_parameter_estimation")
}

save(out, file = "outputs/Bayesian_parameter_estimation/out_mcmc.Rdata")
```

Two sets of posteriors are created, $params and $predict. First, summarize the $predict posteriors and align with empirical data. 
```{r}
temp <- data.frame()
preds1 <- data.frame()
preds2 <- data.frame()
preds <- list()
# Add columns to datlist for each type of output
for(i in 1:length(ids)){
  preds1 <- data.frame()
  preds2 <- data.frame()
  for(j in 1:2){# For types = ACi or AQ, stack
    temp <- data.frame()
    for(k in 1:3){
      sumDF <- coda.fast(out[[i]][[j]][[k]]$predict)[
        grep("pA", row.names(coda.fast(out[[i]][[j]][[k]]$predict))), c(1, 3:5)
      ]
      temp <- rbind.data.frame(temp, sumDF)
    }
    preds1 <- rbind.data.frame(preds1, temp)
  }  
  for(j in 3){
    temp <- data.frame()
    for(k in 1:3){# For types = Aci and AQ, as is
      sumDF <- coda.fast(out[[i]][[j]][[k]]$predict)[
        grep("pA", row.names(coda.fast(out[[i]][[j]][[k]]$predict))), c(1, 3:5)
      ]
      temp <- rbind.data.frame(temp, sumDF)
    }
    preds2 <- temp
  }
  
  # add to existing data as new columns
  preds[[i]] <- rbind.data.frame(cbind.data.frame(datlist[[i]], preds1),
                                 cbind.data.frame(datlist[[i]], preds2))
  preds[[i]]$data <- rep(c("single", "both"), each = nrow(datlist[[i]]))
}

```

Plot predicted and empirical ACi and AQ curves. 
```{r}
# Create folder to save plots
if(dir.exists("outputs/Bayesian_parameter_estimation/plots") == F){
  dir.create("outputs/Bayesian_parameter_estimation/plots")
}

for(i in 1:length(preds)){
  preds[[i]]$data <- factor(preds[[i]]$data, levels = c("single", "both"))
  aciDF <- subset(preds[[i]], type == "ACi" & data == "single" | data == "both")
  fig_aci <- ggplot(aciDF, aes(x = Pci))+
  geom_point(aes(y = A, color = "Observed", shape = type))+
  geom_pointrange(aes(y = mean, ymin = pc2.5, ymax = pc97.5, color = "Predicted"), alpha = 0.5)+
  facet_grid(rep~data)+
  theme_bw(base_size = 12)+
  scale_color_manual(values = c("black", "forestgreen"))
  
  aqDF <- subset(preds[[i]], type == "AQ" & data == "single" | data == "both")
  fig_aq <- ggplot(aqDF, aes(x = Qin))+
  geom_point(aes(y = A, color = "Observed", shape = type))+
  geom_pointrange(aes(y = mean, ymin = pc2.5, ymax = pc97.5, color = "Predicted"), alpha = 0.5)+
  facet_grid(rep~data)+
  theme_bw(base_size = 12)+
  scale_color_manual(values = c("black", "forestgreen"))
  
  # Print out separately
  jpeg(filename = paste0("outputs/Bayesian_parameter_estimation/plots/", ids[i], "_ACi.jpg"),
       height = 4, width = 6, units = "in", res = 600)
  print(fig_aci)
  dev.off()
  
  jpeg(filename = paste0("outputs/Bayesian_parameter_estimation/plots/", ids[i], "_AQ.jpg"),
       height = 4, width = 6, units = "in", res = 600)
  print(fig_aq)
  dev.off()
}
```

Summarize the posteriors into long format labeled with ID, type, rep, and trait.  
```{r}
types <- c("ACi", "AQ", "both")
master <- data.frame()
for(i in 1:length(ids)){
  for(j in 1:length(types)){
    for(k in 1:3){
      sumDF <- coda.fast(out[[i]][[j]][[k]]$params)
      sumDF$trait <- row.names(sumDF)
      sumDF <- subset(sumDF, trait %in% c("vmax0", "r0", "alpha0"))
      sumDF$ID <- rep(ids[i], nrow(sumDF))
      sumDF$type <- rep(types[j], nrow(sumDF))
      sumDF$rep <- rep(paste0("plant_", k), nrow(sumDF))
      master <- rbind.data.frame(master, sumDF)
    }
  }
}
```

Add in traits from Tonantzin's workflow. 
```{r}
# Select vmax, Rd, and AQY for the relevant experiment IDs
params <- read.csv("outputs/parameters_data.csv")
add <- subset(params, select = c(ID, rep, trait, Value, SE, SD))
add <- subset(add, trait %in% c("vmax", "Rd", "AQY") & ID %in% ids)

# Convert traits in master to the same
masterDF <- master %>%
  mutate(trait = case_when(trait == "vmax0" ~ "vmax", 
                           trait == "alpha0" ~ "AQY",
                           trait == "r0" ~ "Rd")) %>%
  left_join(add, by = c("ID", "rep", "trait")) %>%
  select(ID, rep, type, trait, 
         Value, SE, SD, # traits estimated from Tonantzin's workflow
         mean, median, sd, pc2.5, pc97.5) #traits from this analysis

# Create another error columnn that is either SE or SD, for plotting purposes only
masterDF$S2 <- ifelse(is.na(masterDF$SD), masterDF$SE, masterDF$SD)
```

Create 1:1 figures for each trait with Tonantzin's estimates as the baseline. For vmax, error bars are +/- 1 SD (Bayesian) on the x-axis. For AQY and Rd, error bars are +/- 1 SE (frequentist) on the x-axis. All vertical error bars are are +/- 1 SD (Bayesian). 
```{r}
vmaxDF <- subset(masterDF, trait == "vmax")
fig1 <- ggplot(vmaxDF)+
  geom_abline(slope = 1, intercept = 0, col="red")+
  geom_point(aes(x = Value, y = mean))+
  geom_errorbar(aes(x = Value, ymin = mean - SD, ymax = mean + SD))+
  geom_errorbarh(aes(y = mean, xmin = Value - S2, xmax = Value + S2))+
  scale_x_continuous(expression(paste(V[max], " (T)")))+
  scale_y_continuous(expression(paste(V[max], " (fitA)")))+
  facet_wrap(~type)+
  theme_bw(base_size = 14)+
  coord_fixed()
print(fig1)

aqyDF <- subset(masterDF, trait == "AQY")
fig2 <- ggplot(aqyDF)+
  geom_abline(slope = 1, intercept = 0, col="red")+
  geom_point(aes(x = Value, y = mean))+
  geom_errorbar(aes(x = Value, ymin = pc2.5, ymax = pc97.5))+
  geom_errorbarh(aes(y = mean, xmin = Value - S2, xmax = Value + S2))+
  scale_x_continuous(expression(paste(alpha, " (T)")))+
  scale_y_continuous(expression(paste(alpha, " (ftA)")))+
  facet_wrap(~type)+
  theme_bw(base_size = 14)+
  coord_fixed()
print(fig2)

rdDF <- subset(masterDF, trait == "Rd")
fig3 <- ggplot(rdDF)+
  geom_abline(slope = 1, intercept = 0, col="red")+
  geom_point(aes(x = Value, y = mean))+
  geom_errorbar(aes(x = Value, ymin = pc2.5, ymax = pc97.5))+
  geom_errorbarh(aes(y = mean, xmin = Value - S2, xmax = Value + S2))+
  scale_x_continuous(expression(paste(R[d], " (T)")))+
  scale_y_continuous(expression(paste(R[d], " (fitA)")))+
  facet_wrap(~type)+
  theme_bw(base_size = 14)+
  coord_fixed()
print(fig3)

```


#initial test, null model
flux.data = dat
pathway = "C4"
model = NULL
cov.data = NULL

foo <- fitA(flux.data = dat, pathway = "C4")

#test with RE
leaf.model <- list(a.fixed = NULL, a.random = "leaf", 
                V.fixed= NULL, V.random = "leaf",
                k.fixed = NULL, k.random = "leaf",
                n.iter = 5000, match = "fname")

foo <- fitA(flux.data = dat, pathway = "C4", model = leaf.model)


#test with FE
cov.data = read.csv(system.file("extdata", "cov.csv", package = "PEcAn.photosynthesis"))

N.model <- list(a.fixed = NULL, a.random = "leaf", 
                V.fixed= "N", V.random = NULL,
                n.iter = 5000, match = "fname")
foo <- fitA(flux.data = dat, pathway = "C4", model = N.model, cov.data = cov.data)

#test with Danforth data
dat <- read.csv("/home/jessicaguo/sentinel-detection/parameters_pipeline/cleaned_data/ACi/A_Ci_curves_20200304.csv")
#restrict to a single plant
dat2 <- subset(dat, rep == "plant_1")
null.model <- list(V.fixed = NULL, V.random = NULL, 
                          a.fixed = NULL, a.random = NULL, 
                          k.fixed = NULL, k.random = NULL, 
                          n.iter = 5000, match = "rep")
foo <- fitA(flux.data = dat2, model=null.model, pathway = "C4", licor = "6800")

#check output
colnames(foo[[1]][[1]])
s<-summary(foo[[1]])
s$statistics
