---
title: "Plotting relevant parameters between treatments"
author: "Jessica Guo"
output: github_document
urlcolor: blue
editor_options: 
chunk_output_type: console
---

Load necessary packages and functions. 
```{r}
library(ggplot2)
library(data.table)
library(agricolae)
library(dplyr)
library(grid)
```

Load full parameter data. 
```{r}
params <- read.csv("outputs/parameters_data.csv")
```

Subset to include 5 parameters of interest (included in BETY and used by BioCro). 
```{r}
sub <- data.table(subset(params, trait %in% c("vmax", "Rd", "AQY", "theta_lc", "g0BB", "g1BB")))
```

Summarize mean, sd, se, and n by treatment and trait. 
```{r}
sumdf <- sub[,.(mean = mean(Value), 
                  sd = sd(Value), 
                  n = length(Value),
                  se = sd(Value)/sqrt(length(Value))), .(Treatment, trait)]
```

Plot by treatment and trait, adding the sample size to the points as text. 
```{r}
# Order traits by photosynthesis related (top row) and stomatal related (bottom row)
sumdf$trait <- factor(sumdf$trait, levels=c("vmax", "AQY", "Rd", "theta_lc", "g0BB", "g1BB"))
# Order treatments by light level
sumdf$Treatment <- factor(sumdf$Treatment, levels=c("31_22_250","31_31_250", "31_22_450",
                                                    "greenhouse", "outdoor_JollyG", "outdoor_5cm"))
# Add a parsed label colum for trait
sumdf <- sumdf %>% mutate(Trait = case_when(trait == "vmax" ~ "V[max]~(µmol~m^{-2}~s^{-1})",
                                            trait == "AQY" ~ "alpha", 
                                            trait == "Rd" ~ "R[dark]~(µmol~m^{-2}~s^{-1})", 
                                            trait == "theta_lc" ~ "theta", 
                                            trait == "g0BB" ~ "g[0]~(mol~m^{-2}~s^{-1})", 
                                            trait == "g1BB" ~ "g[1]"))
sumdf$Trait <- factor(sumdf$Trait, levels=c("V[max]~(µmol~m^{-2}~s^{-1})", "alpha", "R[dark]~(µmol~m^{-2}~s^{-1})", "theta", "g[0]~(mol~m^{-2}~s^{-1})", "g[1]"))

fig1 <- ggplot(sumdf)+
  geom_errorbar((aes(x = Treatment, ymin = mean - se, ymax = mean + se)), width = 0)+
  geom_point(aes(x = Treatment, y = mean), shape = 15, size = 4)+
  geom_text(aes(x = Treatment, y = mean, label=n), col = "white", size = 3.5, vjust = 0.5, fontface = "bold")+
  facet_wrap(~Trait,scale = "free_x", labeller = label_parsed)+
  scale_x_discrete("Treatment")+
  scale_y_continuous("Mean")+
  theme_bw(base_size = 14)+
  theme()+
  coord_flip()
print(fig1)
```

Calculate ANOVAs and Tukey HSD for each trait by treatment. 
```{r}
models <- list()
hsd <- list()
Traits <- unique(sub$trait)
for(i in 1:length(Traits)){
  df <- subset(sub, trait == Traits[i])
  models[[i]] <- lm(Value ~ Treatment, data = df)
  hsd[[i]] <- HSD.test(models[[i]], "Treatment")$groups
}

# Create table of labels
# Function for last element
substrR <- function(x, n) {substr(x, nchar(x)-n+1, nchar(x))}
substrL <- function(x, n) {substr(x, 1, nchar(x)-n)}
labels <- data.frame(trait = rep(Traits, sub[,.(n = length(unique(Treatment))), by = trait]$n),
                     labs = Reduce(rbind, hsd)$groups,
                     Treatment = ifelse(substrR(rownames(Reduce(rbind, hsd)), 1) %in% 1:5,
                                        substrL(rownames(Reduce(rbind, hsd)), 1),
                                        rownames(Reduce(rbind, hsd))))

# Order traits by photosynthesis related (top row) and stomatal related (bottom row)
labels$trait <- factor(labels$trait, levels=c("vmax", "AQY", "Rd", "theta_lc", "g0BB", "g1BB"))
# Order treatments by light level
labels$Treatment <- factor(labels$Treatment, levels=c("31_22_250","31_31_250", "31_22_450",
                                                    "greenhouse", "outdoor_JollyG", "outdoor_5cm"))

# Add location; need to order labels first
labels <- labels[order(labels$trait, labels$Treatment),]
sumdf <- sumdf[order(sumdf$trait, sumdf$Treatment),]
labels$y = sumdf$mean
#ifelse(is.na(sumdf$se), sumdf$mean, sumdf$mean + sumdf$se)

# Add a parsed label colum for trait
labels <- labels %>% mutate(Trait = case_when(trait == "vmax" ~ "V[max]~(µmol~m^{-2}~s^{-1})",
                                              trait == "AQY" ~ "alpha",
                                              trait == "Rd" ~ "R[dark]~(µmol~m^{-2}~s^{-1})",
                                              trait == "theta_lc" ~ "theta",
                                              trait == "g0BB" ~ "g[0]~(mol~m^{-2}~s^{-1})",
                                              trait == "g1BB" ~ "g[1]"))
labels$Trait <- factor(labels$Trait, levels=c("V[max]~(µmol~m^{-2}~s^{-1})", "alpha", "R[dark]~(µmol~m^{-2}~s^{-1})", "theta", "g[0]~(mol~m^{-2}~s^{-1})", "g[1]"))
```

Plot by treatment and trait, label with Tukey HSD results. 
```{r}
fig2 <- ggplot()+
  geom_errorbar(data = sumdf, (aes(x = Treatment, ymin = mean - se, ymax = mean + se)), width = 0)+
  geom_point(data = sumdf, aes(x = Treatment, y = mean), shape = 15, size = 3)+
  geom_text(data = labels, aes(x = Treatment, y = y, label=labs), size = 3.5, vjust = 1.4, hjust = 0.5)+
  facet_wrap(~Trait, scale = "free_x", labeller = label_parsed)+
  scale_x_discrete("Treatment")+
  scale_y_continuous("Mean")+
  theme_bw(base_size = 14)+
  coord_flip()

#add colors by type (tedious)
g <- ggplot_gtable(ggplot_build(fig2))
strip_both <- which(grepl('strip-', g$layout$name))
fills <- c("darkseagreen", "lightsteelblue1", "lightsteelblue1", "darkseagreen", "darkseagreen", "darkseagreen")
k <- 1
for (i in strip_both) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)
print(fig2)
```