---
title: "Data exploration of psent3"
author: "Jessica Guo"
date: "10/22/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
library(lubridate)
library(tidyr)
library(readxl)
library(udunits2)
```

# psent3 lines

Psent3 is engineered *Setaria* line that responds to mandi by closing stomata, such that leaf temperature increases. So far, only biomass and leaf area measurements have been made on untriggered psent3 lines. Below, we determine how many SLA values are available for psent3, and at what growth conditions. [Here](https://docs.google.com/spreadsheets/d/1MepJgwKb18eg0n50C76S5dhAuaM5cUwy/edit#gid=1245723037) is the summary sheet documenting the psent3 lines.

## Specific Leaf Area

SLA values are calculated from leaf area and leaf dry biomass. Below, the data are read in, filtered, and combined to produce specific leaf area in units of m2/kg. 

First, load in data from the chamber experiments, an Excel sheet entitled "manual-measurements-Darpa_setaria_chambers_experiments.xlsx", which consists of multiple tabs. All leaf areas are available in tab 'total_leaf_area'; biomass for the psent3 lines are found in '8th_biomass_mandi'. 

In addition, we can lump the pSENT3_3.1_3 and the pSENT3_3.8_3 genotypes, as they are both homozygous for the transgene. pSENT3_3.8_22 should be considered separately, as it is the wild type. 

```{r}
# Data from chamber experiments
data_path <- "../../sentinel-detection/data/raw_data/biomass/manual-measurements-Darpa_setaria_chambers_experiments.xlsx"
sheets_names <- excel_sheets(data_path)

# Biomass and leaf area for psent lines, 31_22_430
psent <- read_excel(data_path, sheets_names[4]) %>% 
  filter(grepl("psent3", genotype),
         sample_for == "biomass") %>%
  mutate(tier = case_when(grepl("low", location) ~ "bottom",
                          grepl("bot", location) ~ "bottom",
                          grepl("mid", location) ~ "middle",
                          grepl("top", location) ~ "top"),
         chamber = case_when(grepl("GCH157", location) ~ "GCH157",
                             grepl("GCH158", location) ~ "GCH158"),
         type = case_when(grepl("3.8-22", genotype) ~ "wild-type",
                          grepl("3.8-3", genotype) ~ "psent3",
                          grepl("3.1-3", genotype) ~ "psent3"),
         age = as.numeric(difftime(harvest_time, `sowing date`, "days")),
         ldmc = leaf_fw_g/leaf_dw_g) %>%
  select(plantID, genotype, type, tier, chamber, age, ldmc, height_top_panicle_cm:stem_dw_g)
  
psent %>%
  select(plantID, genotype, type, tier, chamber, age, ldmc, panicle_dw_g:stem_dw_g) %>%
  pivot_longer(ldmc:stem_dw_g, names_to = "trait", values_to = "value") %>%
  filter(trait != "ldmc") %>%
  ggplot(aes(x = trait, y = value, color = type)) +
  geom_point() +
  facet_wrap(~chamber)

la <- read_excel(data_path, sheets_names[2]) %>% 
  filter(grepl("psent3", genotype)) %>%
  rename(leaf_area_cm2 = `total_leaf_area_cm2 (pixels area * factor= 0.000071)`)

sla <- left_join(psent, la) %>%
  mutate(leaf_dw_kg = ud.convert(leaf_dw_g, "g", "kg"), 
         leaf_area_m2 = ud.convert(leaf_area_cm2, "cm2", "m2"),
         sla = leaf_area_m2 / leaf_dw_kg)

ggplot(sla, aes(x = tier, y = sla, col = type, shape = chamber)) +
  geom_point()

m1 <- aov(sla ~ tier, data = sla)
summary(m1)

ggplot(filter(sla, chamber == "GCH158"), aes(x = tier, y = sla, col = type, shape = chamber)) +
  geom_point()
```

The SLA appears to differ strongly by tier, at least witin GCH 157. In GCH 158, both the wild type and the homozygous mutant were grown in the bottom tier, but it appears that variation within the mutants were much larger than variation within the wild-type. 