---
title: "How to Do Regional Runs on DARPA Data Using BioCro Model"
author: "Kristina Riemer, University of Arizona"
output: github_document
urlcolor: blue
---

## Read in libraries

```{r, eval=FALSE}
library(ncdf4)
library(PEcAn.all)
library(PEcAn.BIOCRO)
library(dplyr)
library(ncdf.tools)
library(ggplot2)
library(maps)
library(gganimate)
```

## Read in input files

Past run config file: 
```{r, eval=FALSE}
config <- read.biocro.config("biocro_darpa_results/run/SA-median/config.xml")
```

Data files: 
```{r, eval=FALSE}
metfile <- "biocro_regional_files/champaign.nc"
met_champaign <- nc_open(metfile)
soil_nc <- nc_open("biocro_regional_files/hwsd.nc")
```

## Convert weather data

The weather data is a subset of NARR around Champaign, Illinois. It's currently in the Pecan standard format, which is a netCDF file. The dimensions are 3 latitude, 6 longitude, and 93,504 time steps, which are hourly from 1979 through 2010. There are various weather variables, including temperature, precipitation flux, and solar radiation. 

This is converted to the met format for BioCro. Data for each of the locations (unique latitude and longitude combos) is turned into a flat csv file, with one row per hour for the ~32 years. 
```{r, eval=FALSE}
# Get all locations
lat <- ncvar_get(met_champaign, "latitude")
lon <- ncvar_get(met_champaign, "longitude")
latlon <- expand.grid(lat = lat, lon = lon)

# Get met conversion inputs that apply across all locations
time_vec <- ncvar_get(met_champaign, "time")
time_origin1 <- ncatt_get(met_champaign, "time", "units")
time_origin2 <- gsub( ".*(\\d{4}-\\d{2}-\\d{2}).*", "\\1", time_origin1$value)
start_date <- as.Date(time_origin2) + time_vec[1]
end_date <- as.Date(time_origin2) + time_vec[length(time_vec)] - 1

# Convert met from Pecan to BioCro format for all locations
met_nc <- ncdf4::nc_open(metfile)
point <- 1
for(point in 1:nrow(latlon)){
  met <- load.cfmet(met_nc, lat = latlon$lat[point], lon = latlon$lon[point],
                    start.date = start_date, end.date = end_date)
  biocro_met <- cf2biocro(met)
  biocro_met <- biocro_met %>%
    mutate(RH = case_when(RH >= 1 ~ 0.99999999999999,
                          RH < 1 ~ RH))
  biocro_met_path <- paste0("biocro_regional_files/biocro_met_by_location/biocromet-", 
                            latlon$lat[point], "-", latlon$lon[point], ".2004.csv")
  write.csv(biocro_met, biocro_met_path, row.names = FALSE)
  
}
```

## Run BioCro on each location

BioCro is run on each of the locations weather and soil data to produce daily estimates for a year for biomass, transpiration, etc. 
```{r, eval=FALSE}
biocro_results <- c()
for(point in 1:nrow(latlon)){
  biocro_met_path <- paste0("biocro_regional_files/biocro_met_by_location/biocromet-", 
                            latlon$lat[point], "-", latlon$lon[point])
  biocro_results_all <- run.biocro(latlon$lat[point], latlon$lon[point],
                               metpath = biocro_met_path,
                               soil.nc = soil_nc,
                               config = config)
  biocro_results_daily <- biocro_results_all$daily
  write.csv(biocro_results_daily, paste0("biocro_regional_files/results_by_location/daily-", 
                                         latlon$lat[point], "-", latlon$lon[point], ".csv"))
  biocro_results_daily$lat <- latlon$lat[point]
  biocro_results_daily$lon <- latlon$lon[point]
  biocro_results <- rbind(biocro_results, biocro_results_daily)
  write.csv(biocro_results, "biocro_regional_files/all_biocro_results.csv")
}
```

## Plot estimated biomass

The code below produces a video of a map of total biomass (stem, leaf, grain, rhizome, and root) estimates across time. 
```{r, eval=FALSE}
biocro_results <- biocro_results %>% 
  mutate(date = as.Date(doy, "2003-12-31"), 
         latlon = paste0(lat, lon), 
         total_biomass = Stem + Leaf + Root + Rhizome + Grain)

ggplot(biocro_results, aes(x = date, y = total_biomass, group = latlon, color = latlon)) +
  geom_line() +
  theme(legend.position = "none")

background_map <- map_data("state") %>% 
  filter(region == "illinois")

colfunc <- colorRampPalette(c("darkgreen", "yellow"))
biomass_animation <- ggplot() +
  geom_polygon(data = background_map, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = biocro_results, aes(x = lon, y = lat, fill = total_biomass)) +
  coord_quickmap() +
  scale_fill_gradientn(colors = colfunc(10)) +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

animate(biomass_animation, fps = 100)
anim_save("biomass_animation_champaign.mp4", animation = biomass_animation, 
          path = "./Dropbox/Documents/UofA/Misc/", fps = 100)
```



