---
title: "How to Reproduce BioCro Results for Night Time Temperature Experiments"
author: "Kristina Riemer, University of Arizona"
output: github_document
urlcolor: blue
---

For these experiments, **Setaria** were grown under two experimental treatments. The control plants were at 31$^\circ$C during the day and 22$^\circ$C at night, while high night time temp treatment was 31$^\circ$C at all times. 

Biomass for these two treatments are estimated using BioCro version 0.95. 

# Section 1: BioCro Run for Control

1. Access RStudio in a browser by navigating to [welsch.cyverse.org:8787/](welsch.cyverse.org:8787/). Log in using your username and password. 

2. Create a new folder called `biocro_temp_exps_files`. 

3. Create new XML called `pecan.biocro.darpa.temp.exps.xml`

```{r, eval=FALSE}
<pecan>
  <outdir>biocro_temp_exps_results</outdir>

  <database>
    <bety>
      <driver>PostgreSQL</driver>
      <user>bety</user>
      <password>bety</password>
      <host>postgres</host>
      <dbname>bety</dbname>
      <write>FALSE</write>
    </bety>
    <dbfiles>biocro_temp_exps_results/dbfiles</dbfiles>
  </database>

  <pfts>
    <pft>
      <name>SetariaWT_ME034</name>
    </pft>
  </pfts>

  <ensemble>
    <size>5</size>
    <variable>TotLivBiom</variable>
  </ensemble>

  <meta.analysis>
    <iter>3000</iter>
    <random.effects>TRUE</random.effects>
    <threshold>1.2</threshold>
    <update>AUTO</update>
  </meta.analysis>

  <sensitivity.analysis>
    <variable>TotLivBiom</variable>
  </sensitivity.analysis>

  <model>
    <type>BIOCRO</type>
    <binary>/home/kristinariemer/pecan/models/biocro/inst/biocro.Rscript</binary>
    <revision>0.95</revision>
  </model>

  <run>
    <site>
      <id>9000000004</id>
    </site>
    <inputs>
      <met>
        <output>BIOCRO</output>
        <path>/home/kristinariemer/biocro_temp_exps_files/danforth-chamber</path>
      </met>
    </inputs>
      <start.date>2019/01/01</start.date>
      <end.date>2019/12/30</end.date>
    <host>
      <name>localhost</name>
    </host>
  </run>
</pecan>
```

4. Create new `workflow.R`

```{r, eval=FALSE}
#!/usr/bin/env Rscript
#-------------------------------------------------------------------------------
# Copyright (c) 2012 University of Illinois, NCSA.
# All rights reserved. This program and the accompanying materials
# are made available under the terms of the 
# University of Illinois/NCSA Open Source License
# which accompanies this distribution, and is available at
# http://opensource.ncsa.illinois.edu/license.html
#-------------------------------------------------------------------------------

# ----------------------------------------------------------------------
# Load required libraries
# ----------------------------------------------------------------------
library(PEcAn.all)
library(PEcAn.utils)
library(RCurl)

# make sure always to call status.end
options(warn=1)
options(error=quote({
  PEcAn.utils::status.end("ERROR")
  PEcAn.remote::kill.tunnel(settings)
  if (!interactive()) {
    q(status = 1)
  }
}))

#options(warning.expression=status.end("ERROR"))


# ----------------------------------------------------------------------
# PEcAn Workflow
# ----------------------------------------------------------------------
# Open and read in settings file for PEcAn run.
args <- commandArgs(trailingOnly = TRUE)
if (is.na(args[1])){
  settings <- PEcAn.settings::read.settings("pecan.xml") 
} else {
  settings.file <- args[1]
  settings <- PEcAn.settings::read.settings(settings.file)
}

# Check for additional modules that will require adding settings
if("benchmarking" %in% names(settings)){
  library(PEcAn.benchmark)
  settings <- papply(settings, read_settings_BRR)
}

if("sitegroup" %in% names(settings)){
  if(is.null(settings$sitegroup$nSite)){
    settings <- PEcAn.settings::createSitegroupMultiSettings(settings, 
                                                             sitegroupId = settings$sitegroup$id)
  } else {
    settings <- PEcAn.settings::createSitegroupMultiSettings(settings, 
                                                             sitegroupId = settings$sitegroup$id,
                                                             nSite = settings$sitegroup$nSite)
  }
  settings$sitegroup <- NULL ## zero out so don't expand a second time if re-reading
}

# Update/fix/check settings. Will only run the first time it's called, unless force=TRUE
settings <- PEcAn.settings::prepare.settings(settings, force = FALSE)

# Write pecan.CHECKED.xml
PEcAn.settings::write.settings(settings, outputfile = "pecan.CHECKED.xml")

# start from scratch if no continue is passed in
statusFile <- file.path(settings$outdir, "STATUS")
if (length(which(commandArgs() == "--continue")) == 0 && file.exists(statusFile)) {
  file.remove(statusFile)
}

# Do conversions
settings <- PEcAn.workflow::do_conversions(settings)

# Query the trait database for data and priors
if (PEcAn.utils::status.check("TRAIT") == 0){
  PEcAn.utils::status.start("TRAIT")
  settings <- PEcAn.workflow::runModule.get.trait.data(settings)
  PEcAn.settings::write.settings(settings, outputfile='pecan.TRAIT.xml')
  PEcAn.utils::status.end()
} else if (file.exists(file.path(settings$outdir, 'pecan.TRAIT.xml'))) {
  settings <- PEcAn.settings::read.settings(file.path(settings$outdir, 'pecan.TRAIT.xml'))
}

# Run the PEcAn meta.analysis
if(!is.null(settings$meta.analysis)) {
  if (PEcAn.utils::status.check("META") == 0){
    PEcAn.utils::status.start("META")
    PEcAn.MA::runModule.run.meta.analysis(settings)
    PEcAn.utils::status.end()
  }
}

# Write model specific configs
if (PEcAn.utils::status.check("CONFIG") == 0){
  PEcAn.utils::status.start("CONFIG")
  settings <- PEcAn.workflow::runModule.run.write.configs(settings)
  PEcAn.settings::write.settings(settings, outputfile='pecan.CONFIGS.xml')
  PEcAn.utils::status.end()
} else if (file.exists(file.path(settings$outdir, 'pecan.CONFIGS.xml'))) {
  settings <- PEcAn.settings::read.settings(file.path(settings$outdir, 'pecan.CONFIGS.xml'))
}

if ((length(which(commandArgs() == "--advanced")) != 0) && 
    (PEcAn.utils::status.check("ADVANCED") == 0)) {
  PEcAn.utils::status.start("ADVANCED")
  q();
}

# Start ecosystem model runs
if (PEcAn.utils::status.check("MODEL") == 0) {
  PEcAn.utils::status.start("MODEL")
  PEcAn.remote::runModule.start.model.runs(settings, stop.on.error = FALSE)
  PEcAn.utils::status.end()
}

# Get results of model runs
if (PEcAn.utils::status.check("OUTPUT") == 0) {
  PEcAn.utils::status.start("OUTPUT")
  runModule.get.results(settings)
  PEcAn.utils::status.end()
}

# Run ensemble analysis on model output.
if ('ensemble' %in% names(settings) & PEcAn.utils::status.check("ENSEMBLE") == 0) {
  PEcAn.utils::status.start("ENSEMBLE")
  runModule.run.ensemble.analysis(settings, TRUE)
  PEcAn.utils::status.end()
}

# Run sensitivity analysis and variance decomposition on model output
if ('sensitivity.analysis' %in% names(settings) & PEcAn.utils::status.check("SENSITIVITY") == 0) {
  PEcAn.utils::status.start("SENSITIVITY")
  runModule.run.sensitivity.analysis(settings)
  PEcAn.utils::status.end()
}

# Run parameter data assimilation
if ('assim.batch' %in% names(settings)) {
  if (PEcAn.utils::status.check("PDA") == 0) {
    PEcAn.utils::status.start("PDA")
    settings <- PEcAn.assim.batch::runModule.assim.batch(settings)
    PEcAn.utils::status.end()
  }
}

# Run state data assimilation
if ('state.data.assimilation' %in% names(settings)) {
  if (PEcAn.utils::status.check("SDA") == 0) {
    PEcAn.utils::status.start("SDA")
    settings <- sda.enfk(settings)
    PEcAn.utils::status.end()
  }
}

# Run benchmarking
if("benchmarking" %in% names(settings) & "benchmark" %in% names(settings$benchmarking)){
  PEcAn.utils::status.start("BENCHMARKING")
  results <- papply(settings, function(x) calc_benchmark(x, bety))
  PEcAn.utils::status.end()
}

# Pecan workflow complete
if (PEcAn.utils::status.check("FINISHED") == 0) {
  PEcAn.utils::status.start("FINISHED")
  PEcAn.remote::kill.tunnel(settings)
  db.query(paste("UPDATE workflows SET finished_at=NOW() WHERE id=",
                 settings$workflow$id, "AND finished_at IS NULL"),
           params = settings$database$bety)
  
  # Send email if configured
  if (!is.null(settings$email) && !is.null(settings$email$to) && (settings$email$to != "")) {
    sendmail(settings$email$from, settings$email$to,
             paste0("Workflow has finished executing at ", base::date()),
             paste0("You can find the results on ", settings$email$url))
  }
  PEcAn.utils::status.end()
}

db.print.connections()
print("---------- PEcAn Workflow Complete ----------")
```

5. Generate weather data `danforth-chamber.2019.csv` with this R script. 

```{r, eval=FALSE}
hightempchamber_weather <- data.frame(year = rep(2019, 8760), 
                                      doy = rep(1:365, each = 24), 
                                      hour = rep(seq(0, 23), 365), 
                                      SolarR = rep(c(0, 936), times = 365, each = 12),
                                      Temp = rep(31, times = 365 * 24), 
                                      RH = rep(55.5 / 100,  times = 365 * 24), 
                                      WS = rep(0, times = 365 * 24), 
                                      precip = rep(c(0.000462963, rep(0, 23)), 365))

write.csv(hightempchamber_weather, "biocro_temp_exps_files/danforth-chamber.2019.csv", 
          row.names = FALSE)
```

6. To run model for control treatment. 

```{bash, eval=FALSE}
biocro_temp_exps_files/workflow.R --settings biocro_temp_exps_files/pecan.biocro.darpa.temp.exps.xml
```

# Section 2: BioCro Run for High Night Temp

