---
title: "How to Optimze Biomass Estimation for DARPA Data Using BioCro Model"
author: "Kristina Riemer, University of Arizona"
output: github_document
urlcolor: blue
---

### Walkthrough for optimizing using `DEoptim`

Optimizing biomass coefficients so that biomass estimates from BioCro as 
closely as possible match the measured biomass values. Parameters of interest are phenoParms, with thermal times and rhizome coefficients being fixed and leaf, stem, root, and grain coefficients being optimized across. This returns values of biomass for each stage and plant part, but the value being minimized is the difference between measured biomass values and biomass estimated by BioCro. It seems like we can optimized across all stages at once within our objective function. 

First read in all the necessary input data: 

* BioCro config file
* weather file created for biomass partitioning
* biomass file from Setaria experiments

```{r}
library(dplyr)

rundir <- getwd()

OpBioGro_weather <- read.csv("biocro_opt_darpa_files/OpBioGro_weather.csv")
colnames(OpBioGro_weather) <- c("year", "doy", "hour", "Solar", "Temp", "RH", "WS", "precip")
readr::write_csv(OpBioGro_weather, path = file.path(rundir, 'SA-median.2019.csv'))
WetDat <- OpBioGro_weather

config <- PEcAn.BIOCRO::read.biocro.config(file.path(rundir, "biocro_opt_darpa_files/config.xml"))

lat <- as.numeric(config$location$latitude)
lon <- as.numeric(config$location$longitude)
config$run$start.date <- "2019/01/01"
config$run$end.date <- "2019/12/01"
start.date <- lubridate::date(config$run$start.date)
end.date   <- lubridate::date(config$run$end.date)
genus <- config$pft$type$genus
years <- lubridate::year(start.date):lubridate::year(end.date)
l2n <- function(x) lapply(x, as.numeric)
day1 <- 1
dayn <- 100

OpBioGro_biomass <- read.csv("biocro_opt_darpa_files/OpBioGro_biomass.csv")
```

We want biomass estimates from `BioGro` (i.e., ttt) to be as close to the actual biomass values as possible (i.e., OpBioGro_biomass). Will be optimizing over the parameters, which are the biomass coefficients. 

TODO:

* ~~Remove thermal times from biomass tables~~
* (later) figure out better metric for comparing biomass values
  * ~~sum of absolute value (because sign doesn't matter, magnitude of difference does)~~
  * log transform
  * (ttt - OpBioGro_biomass) / OpBioGro_biomass
* ~~upper and lower bounds are, initially, each between 0 and 1~~
  * Can use more restricted bounds, base on BioCro defaults? 
* ~~additional param constraints: parameter values for each stage should sum to one (note: this MUST happen because `BioCro` will not use pheno params that do not equal one)~~
  * ~~fnMap argument is only for number of parameters constraints?~~ 
  * ~~include something like the below, see `?DEoptim` for another example~~
function(x){
 if(x[1:5] == 1){
 ifelse(x[6:10]){
 ...
 }
}

  * ~~another option for this sum to one constraint is something like the following for each stage as the first part of `opfn`~~

```{r, eval=FALSE}
z <- c(0.1, 0.2, 0.15)
z[1:3]
sum(z[1:3])
z[1:3]/sum(z[1:3])
sum(z[1:3]/sum(z[1:3]))
```

* ~~keep thermal times the same, can set them in the phenoControl line, and set rhizome values == 0~~

* ~~phenoParms = vector of 25 values~~
  * ~~split up parameters into the ones that will vary and the ones that will not. op parms = 19 values, thermal = 6, rhizome = 6~~
  * ~~reassemble within opfn function~~
* ~~Biomass measurements need to be in correct units (Mg/ha)~~

First create objective function `opfn`, and include parameter values to test if this function by itself works. 
```{r}
library(BioCro)

opfn <- function(optimizingParms, thermaltimeParms, rhizomeParms){ 
  optimizingParms[1:3] <- optimizingParms[1:3]/sum(optimizingParms[1:3])
  optimizingParms[4:6] <- optimizingParms[4:6]/sum(optimizingParms[4:6])
  optimizingParms[7:9] <- optimizingParms[7:9]/sum(optimizingParms[7:9])
  optimizingParms[10:12] <- optimizingParms[10:12]/sum(optimizingParms[10:12])
  optimizingParms[13:15] <- optimizingParms[13:15]/sum(optimizingParms[13:15])
  optimizingParms[16:19] <- optimizingParms[16:19]/sum(optimizingParms[16:19])
  params_list <- phenoParms(tp1 = thermaltimeParms[1], 
                           tp2 = thermaltimeParms[2], 
                           tp3 = thermaltimeParms[3], 
                           tp4 = thermaltimeParms[4], 
                           tp5 = thermaltimeParms[5], 
                           tp6 = thermaltimeParms[6], 
                           kStem1 = optimizingParms[1], 
                           kStem2 = optimizingParms[4], 
                           kStem3 = optimizingParms[7], 
                           kStem4 = optimizingParms[10], 
                           kStem5 = optimizingParms[13], 
                           kStem6 = optimizingParms[16], 
                           kLeaf1 = optimizingParms[2], 
                           kLeaf2 = optimizingParms[5], 
                           kLeaf3 = optimizingParms[8], 
                           kLeaf4 = optimizingParms[11], 
                           kLeaf5 = optimizingParms[14], 
                           kLeaf6 = optimizingParms[17], 
                           kRoot1 = optimizingParms[3], 
                           kRoot2 = optimizingParms[6], 
                           kRoot3 = optimizingParms[9], 
                           kRoot4 = optimizingParms[12], 
                           kRoot5 = optimizingParms[15], 
                           kRoot6 = optimizingParms[18], 
                           kRhizome1 = rhizomeParms[1], 
                           kRhizome2 = rhizomeParms[2], 
                           kRhizome3 = rhizomeParms[3], 
                           kRhizome4 = rhizomeParms[4], 
                           kRhizome5 = rhizomeParms[5], 
                           kRhizome6 = rhizomeParms[6], 
                           kGrain6 = optimizingParms[19])
  #print(params_list)
  t <- BioCro::BioGro(
    WetDat = WetDat,
    day1 = day1,
    dayn = dayn,
    iRhizome = 0.001, 
    iLeaf = 0.001, 
    iStem = 0.001, 
    iRoot = 0.001, 
    soilControl = l2n(config$pft$soilControl),
    canopyControl = l2n(config$pft$canopyControl),
    phenoControl = l2n(params_list),
    seneControl = l2n(config$pft$seneControl),
    photoControl = l2n(config$pft$photoParms))

  tt <- data.frame(ThermalT = t$ThermalT, 
                   Stem = t$Stem,
                   Leaf = t$Leaf,
                   Root = t$Root,
                   Rhizome = t$Rhizome,
                   Grain = t$Grain,
                   LAI = t$LAI)
  ttt <- tt %>% 
    filter(round(tt$ThermalT) %in% round(OpBioGro_biomass$ThermalT))
  bio_ests <- select(ttt, -ThermalT)
  bio_meas <- select(OpBioGro_biomass, -ThermalT)
  diff <- abs(bio_ests - bio_meas)
  
  bio_meas_sum <- bio_meas %>% 
    mutate(stage = rownames(.)) %>% 
    rowwise() %>% 
    mutate(total_bio = sum(Stem:Grain)) %>% 
    select(stage, total_bio)
  
  diff_norm <- diff %>% 
    mutate(stage = rownames(.)) %>% 
    left_join(., bio_meas_sum, by = c("stage")) %>% 
    mutate(Stem_norm = Stem / total_bio, 
         Leaf_norm = Leaf / total_bio, 
         Root_norm = Root / total_bio) %>% 
    select(Stem_norm, Leaf_norm, Root_norm, Rhizome, Grain, LAI) %>% 
    rename(Stem = Stem_norm, Leaf = Leaf_norm, Root = Root_norm)
  
  return(sum(diff_norm))
}

# Check function itself by running with example parameter values
optimizingParms_check <- c(rep(c(0.2, 0.3, 0.5), 6), 0)
thermaltimeParms_check <- c(config$pft$phenoParms$tp1, config$pft$phenoParms$tp2,
                      config$pft$phenoParms$tp3, config$pft$phenoParms$tp4,
                      config$pft$phenoParms$tp5, config$pft$phenoParms$tp6)
rhizomeParms_check <- rep(0, 6)
opfn(optimizingParms_check, thermaltimeParms_check, rhizomeParms_check)
```

Run objective function through optimization with `DEoptim`, setting the upper and lower bounds for the varying parameters to 0 and 1 and providing values for the non-varying parameters thermal time and rhizome coefficients. Reducing itermax reduces time to run `DEoptim`, but the resulting best value doesn't change much. 
```{r, results='hide', message=FALSE}
library(DEoptim)

thermaltimevals <- c(config$pft$phenoParms$tp1, config$pft$phenoParms$tp2, 
                     config$pft$phenoParms$tp3, config$pft$phenoParms$tp4, 
                     config$pft$phenoParms$tp5, config$pft$phenoParms$tp6)
rhizomevals <- rep(0, 6)

opt_results <- DEoptim(fn = opfn, lower = rep(0, 19), upper = rep(1, 19), thermaltimeParms = thermaltimevals, rhizomeParms = rhizomevals, control = DEoptim.control(itermax = 2))
```

Test that the resulting parameter values produce the same biomass estimates. `parms_results` are the actual parameters that we would use because they've been adjusted for the sum to 1 constraint. 
```{r}
library(ggplot2)

parms_results <- as.vector(opt_results$optim$bestmem)
parms_results[1:3] <- parms_results[1:3]/sum(parms_results[1:3])
parms_results[4:6] <- parms_results[4:6]/sum(parms_results[4:6])
parms_results[7:9] <- parms_results[7:9]/sum(parms_results[7:9])
parms_results[10:12] <- parms_results[10:12]/sum(parms_results[10:12])
parms_results[13:15] <- parms_results[13:15]/sum(parms_results[13:15])
parms_results[16:19] <- parms_results[16:19]/sum(parms_results[16:19])
optimalParms <- phenoParms(tp1 = thermaltimevals[1], 
                                 tp2 = thermaltimevals[2], 
                                 tp3 = thermaltimevals[3], 
                                 tp4 = thermaltimevals[4], 
                                 tp5 = thermaltimevals[5], 
                                 tp6 = thermaltimevals[6], 
                                 kStem1 = parms_results[1], 
                                 kStem2 = parms_results[4], 
                                 kStem3 = parms_results[7], 
                                 kStem4 = parms_results[10], 
                                 kStem5 = parms_results[13], 
                                 kStem6 = parms_results[16], 
                                 kLeaf1 = parms_results[2], 
                                 kLeaf2 = parms_results[5], 
                                 kLeaf3 = parms_results[8], 
                                 kLeaf4 = parms_results[11], 
                                 kLeaf5 = parms_results[14], 
                                 kLeaf6 = parms_results[17], 
                                 kRoot1 = parms_results[3], 
                                 kRoot2 = parms_results[6], 
                                 kRoot3 = parms_results[9], 
                                 kRoot4 = parms_results[12], 
                                 kRoot5 = parms_results[15], 
                                 kRoot6 = parms_results[18], 
                                 kRhizome1 = rhizomevals[1], 
                                 kRhizome2 = rhizomevals[2], 
                                 kRhizome3 = rhizomevals[3], 
                                 kRhizome4 = rhizomevals[4], 
                                 kRhizome5 = rhizomevals[5], 
                                 kRhizome6 = rhizomevals[6], 
                                 kGrain6 = parms_results[19])

results_test <- BioCro::BioGro(
    WetDat = WetDat,
    day1 = day1,
    dayn = dayn,
    iRhizome = 0.001, 
    iLeaf = 0.001, 
    iStem = 0.001, 
    iRoot = 0.001, 
    soilControl = l2n(config$pft$soilControl),
    canopyControl = l2n(config$pft$canopyControl),
    phenoControl = l2n(optimalParms),
    seneControl = l2n(config$pft$seneControl),
    photoControl = l2n(config$pft$photoParms))

results_test2 <- data.frame(ThermalT = results_test$ThermalT, 
                 Stem = results_test$Stem,
                 Leaf = results_test$Leaf,
                 Root = results_test$Root,
                 Rhizome = results_test$Rhizome,
                 Grain = results_test$Grain,
                 LAI = results_test$LAI)
results_test3 <- results_test2 %>% 
  filter(round(results_test2$ThermalT) %in% round(OpBioGro_biomass$ThermalT))
diff <- sum(abs(results_test3 - OpBioGro_biomass))

plot(results_test)

biomass_meas_plot <- OpBioGro_biomass %>% 
  tidyr::pivot_longer(Stem:Grain)
biomass_ests_plot <- results_test2 %>% 
  tidyr::pivot_longer(Stem:Grain)
ggplot(biomass_meas_plot, aes(ThermalT, value, color = name)) +
  geom_point() +
  geom_line(data = biomass_ests_plot, aes(ThermalT, value, color = name)) +
  xlim(c(0, 2000)) +
  labs(x = "Thermal Time", y = "Biomass (Ma/ha)", color = "Plant Part") +
  theme_classic()
```

### DEoptim examples

#### Simple

Has ten parameters (from n), which is also number of values in upper and lower bounds vector? 

`ans$optim$bestmem` shows best ten parameter values, and `ans$optim$bestval` shows the resulting fval for those parameters values. 

Minimizing fval

```{r, eval=FALSE}
genrose.f <- function(x){
  n <- length(x)
  fval <- 1.0 + sum (100 * (x[1:(n-1)]^2 - x[2:n])^2 + (x[2:n] - 1)^2)
  return(fval)
}
n <- 10
ans <- DEoptim(fn=genrose.f, lower=rep(-5, n), upper=rep(5, n),
                control=list(NP=100, itermax=4000,trace=FALSE))
```

#### More complicated example

This includes varying and non-varying parameters. This runs `DEoptim` multiple times in a loop for different non-varying params. 

```{r, eval=FALSE}
library(NMOF)
# Define objective function
error_heston <- function(x, P, S, K, tau, r, q) {
  v0 <- x[1]
  vT <- x[2]
  rho <- x[3]
  k <- x[4]
  sigma <- x[5]
  error <- abs(P - callHestoncf(S, K, tau, r, q, v0, vT, rho, k, sigma))
  return(error)
}

# Only need to set bounds for varying parameters
lower <- c(1e-4, 1e-4, -1, 1e-4, 1e-4)  
upper <- c(  10, 10,    1,   10,   10)

# Real option price
C0116_P <- c(1328.70, 1316.05, 1333.05, 1337.90, 1344.80)
# Strike price
C0116_K <- c(500, 500, 500, 500, 500)
# Time to maturity in years
C0116_T_t <- c(1.660274, 1.657534, 1.654795, 1.652055, 1.649315)
# Interest rate percentage
C0116_r <- c(0.08, 0.08, 0.09, 0.09, 0.09)
# Dividend yield percentage
C0116_DY <- c(2.07, 2.09, 2.07, 2.07, 2.06) / 100
# Price underlying
C0116_SP_500_P <- c(1885.08, 1872.83, 1888.03, 1892.49, 1900.53)

i <- 1
outDEoptim <- DEoptim(error_heston, lower, upper,
  DEoptim.control(VTR=0, itermax=100), P = C0116_P[i], S = C0116_SP_500_P[i],
  K = C0116_K[i], tau = C0116_T_t[i], r = C0116_r[i], q = C0116_DY[i])
```


### MCMC approach

BayesianSetup = model info and model parameter priors
model = BioGro
parameters = (varying) leaf, stem, root coeffs (& non-varying?)

provide prior for each param, can have upper and lower boundaries and distribution

likelihood = model

Example has three parameters with uniform distributions bounded at -10 and 10. 

```{r, eval=FALSE}
library(BayesianTools)

ll_ex <- generateTestDensityMultiNormal(sigma = "no correlation")
bayesianSetup_ex <- createBayesianSetup(likelihood = ll_ex, lower = rep(-10, 3), upper = rep(10, 3))

iter_ex = 10000
settings_ex = list(iterations = iter_ex, nrChains = 3, message = FALSE)
out <- runMCMC(bayesianSetup = bayesianSetup_ex, sampler = "Metropolis", settings = settings_ex)

plot(out)
getSample(out, start = 100, end = NULL, thin = 5, whichParameters = 1:2)
```

Possible ll for our case: 
```{r, eval=FALSE}
sum(log(dnorm(observed_biomasses-predicted_biomasses,0,1)))
```

```{r, eval=FALSE}
ll <- function(x) sum(dnorm(x, log = TRUE))
test <- createBayesianSetup(ll, prior = NULL, priorSampler = NULL, lower = -10, upper = 10)
str(test)
test$prior$density(0)
test$likelihood$density(c(1,1))
test_run <- runMCMC(test)
plot(test_run)
```

prior distribution + likelihood distribution = posterior distribution
MCMC is for estimating posterior when it can't be calculated directly (e.g., from unusual prior and likelihood dists)
