---
title: "How to Do Regional Runs on DARPA Data Using BioCro Model"
author: "Author: Kristina Riemer"
output: github_document
urlcolor: blue
---

## Read in libraries

```{r}
library(ncdf4)
library(PEcAn.all)
library(PEcAn.BIOCRO)
library(dplyr)
library(ncdf.tools)
library(ggplot2)
library(maps)
library(speciesgeocodeR)
library(rgdal)
library(maptools)
library(gganimate)
```

## Read in input files

The config file here is based on [the default *Setaria* file](https://github.com/PecanProject/pecan/blob/ac5aef2e755a8961591196428753eb184fb4c6cb/models/biocro/inst/extdata/defaults/setaria.xml) used for PEcAn. This one was generated from a run on the Sentinel project data. 
```{r}
config <- read.biocro.config("biocro_regional_darpa_files/config.xml")
```

A weather data file, such as that read in below, is required to run BioCro. The weather data comes from [the North American Regional Reanalysis (NARR)](https://www.esrl.noaa.gov/psd/data/gridded/data.narr.html) global gridded dataset provided by NOAA. We are using a subset of NARR in the eastern part of Illinois, near Chapaign. This subset is in the Pecan standard format, which is a netCDF file. The dimensions are latitude x longitude x time (3 x 6 x 93,504). These time steps are every three hours from 1979 through 2010. There are 10 weather variables, including temperature, precipitation flux, and solar radiation. 

We also read in a soil file, which is not necessary but is useful. These data are from [the harmonized world soil database](http://www.fao.org/soils-portal/soil-survey/soil-maps-and-databases/harmonized-world-soil-database-v12/en/). 
```{r}
metfile <- "biocro_regional_darpa_files/chiruss.nc"
met_chiruss <- nc_open(metfile)
soil_nc <- nc_open("biocro_regional_darpa_files/hwsd.nc")
```

## Convert weather data

The weather dataset has to be converted from netCDF file format to the format required for BioCro. Data for each of the locations (unique latitude and longitude combinations) is turned into a flat csv file, with one row per hour for the time range of the weather dataset (approximately 32 years). 
```{r, eval=FALSE}
# Get all locations
lat <- ncvar_get(met_chiruss, "latitude")
lon <- ncvar_get(met_chiruss, "longitude")
latlon <- expand.grid(lat = lat, lon = lon)

# Get met conversion inputs that apply across all locations
time_vec <- ncvar_get(met_chiruss, "time")
time_origin1 <- ncatt_get(met_chiruss, "time", "units")
time_origin2 <- gsub( ".*(\\d{4}-\\d{2}-\\d{2}).*", "\\1", time_origin1$value)
start_date <- as.Date(time_origin2) + time_vec[1]
end_date <- as.Date(time_origin2) + time_vec[length(time_vec)] - 1

# Convert met from Pecan to BioCro format for all locations
dir.create("biocro_regional_darpa_files/biocro_met_by_location/")
met_nc <- ncdf4::nc_open(metfile)
point <- 1
biocro_met_locations <- c()
for(point in 1:nrow(latlon)){
  met <- load.cfmet(met_nc, lat = latlon$lat[point], lon = latlon$lon[point],
                    start.date = start_date, end.date = end_date)
  met_hourly <- cfmet.downscale.time(met, output.dt = 1)
  biocro_met <- cf2biocro(met_hourly)
  biocro_met <- biocro_met %>%
    mutate(RH = case_when(RH >= 1 ~ 0.99999999999999,
                          RH < 1 ~ RH))
  biocro_met_location <- biocro_met %>% 
    mutate(latitude = latlon$lat[point], 
           longitude = latlon$lon[point])
  biocro_met_locations <- rbind(biocro_met_locations, biocro_met_location)
  biocro_met_path <- paste0("biocro_regional_darpa_files/biocro_met_by_location/biocromet-", 
                            latlon$lat[point], "-", latlon$lon[point], ".2010.csv")
  write.csv(biocro_met, biocro_met_path, row.names = FALSE)
}
```

We plot the first day of all the weather data variables to examine the daily patterns in them. 
```{r, eval=FALSE}
biocro_met_plot <- biocro_met_locations %>% 
  tidyr::unite(latitudelongitude, latitude:longitude) %>% 
  mutate(date = as.Date(paste0(year, "-01-01")) - 1 + doy, 
         time = case_when(
           hour < 10 ~ paste0("0", hour, ":00"), 
           hour >= 10 ~ paste0(hour, ":00")
         ), 
         datetime_gmt = as.POSIXct(paste(date, time), tz = "GMT"), 
         datetime = as.POSIXct(format(datetime_gmt, tz = "America/Chicago")), 
         date = as.Date(datetime, tz = "America/Chicago"), 
         year_plot = format(date, "%Y"), 
         latitudelongitude = as.factor(latitudelongitude)) 

biocro_met_plot_year <- biocro_met_plot %>% 
  filter(year_plot == 1995 & latitudelongitude == "40.375_-87.875") %>%
  select(datetime, latitudelongitude, solar:precip) %>% 
  tidyr::gather(weather_var, weather_value, solar:precip)
  
ggplot(biocro_met_plot_year, aes(x = datetime, y = weather_value, color = latitudelongitude)) +
  geom_line() +
  facet_wrap(~weather_var, scales = "free_y")

biocro_met_plot_day <- biocro_met_plot %>% 
  filter(date == as.Date("1979-06-01")) %>% 
  select(datetime, latitudelongitude, solar:precip) %>% 
  tidyr::gather(weather_var, weather_value, solar:precip)

ggplot(biocro_met_plot_day, aes(x = datetime, y = weather_value, color = latitudelongitude)) +
  geom_line() +
  facet_wrap(~weather_var, scales = "free_y")
```

## Run BioCro on each location

The BioCro model is then run on weather and soil data for each of the locations, using *Setaria* input data. This produces hourly, daily, and yearly estimates for biomass, transpiration, etc. for the time range we specified in the config file. We are using only the daily values. 
```{r, eval=FALSE}
dir.create("biocro_regional_darpa_files/results_by_location/")
biocro_results <- c()
for(point in 1:nrow(latlon)){
  biocro_met_path <- paste0("biocro_regional_darpa_files/biocro_met_by_location/biocromet-", 
                            latlon$lat[point], "-", latlon$lon[point])
  biocro_results_all <- run.biocro(latlon$lat[point], latlon$lon[point],
                               metpath = biocro_met_path,
                               soil.nc = soil_nc,
                               config = config)
  biocro_results_daily <- biocro_results_all$daily
  write.csv(biocro_results_daily, paste0("biocro_regional_darpa_files/results_by_location/daily-", 
                                         latlon$lat[point], "-", latlon$lon[point], ".csv"))
  biocro_results_daily$lat <- latlon$lat[point]
  biocro_results_daily$lon <- latlon$lon[point]
  biocro_results <- rbind(biocro_results, biocro_results_daily)
  write.csv(biocro_results, "biocro_regional_darpa_files/all_biocro_results.csv")
}
```

## Plot estimated biomass

The results for total biomass (the sum of stem, leaf, grain, rhizome, and root biomasses) are then visualized. 
```{r}
biocro_results <- read.csv("biocro_regional_darpa_files/chiruss_results/all_biocro_results.csv")
biocro_results <- biocro_results %>% 
  mutate(date = as.Date(doy, "2009-12-31"), 
         latlon = paste0(lat, lon), 
         total_biomass = Stem + Leaf + Root + Rhizome + Grain)

ggplot(biocro_results, aes(x = date, y = total_biomass, group = latlon, color = latlon)) +
  geom_line() +
  theme(legend.position = "none")
```

This map shows where the locations are in the two countries
```{r}
overall_map <- map_data("world") %>% 
  filter(region == "China" | region == "Russia")

ggplot() +
  geom_polygon(data = overall_map, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = biocro_results, aes(x = lon, y = lat), fill = "red") +
  coord_quickmap()
```

This map shows the locations closer up with the nearby ecoregions plotted in various colors. 
```{r}
ecoregions_background <- map_data("world") %>% 
  filter(region == "China" | region == "Russia") 

ecoregions <- readOGR(file.path("biocro_regional_darpa_files", 
                                                "WWF_ecoregions", "official", 
                                                "wwf_terr_ecos.shp"))
ecoregions@data$id <- rownames(ecoregions@data)
ecoregions_points <- fortify(ecoregions, region = "id")
ecoregions_df <- plyr::join(ecoregions_points, ecoregions@data, by = "id")
ecoregions_df_chiruss <- ecoregions_df %>% 
   filter(lat > 20 & lat < 80,
          long > 60 & long < 130)

ggplot() +
  geom_polygon(data = ecoregions_df_chiruss, aes(long, lat, group = group, fill = ECO_NAME)) +
  geom_polygon(data = ecoregions_background, aes(x = long, y = lat, group = group), fill = NA, color = "black") +
  geom_raster(data = biocro_results, aes(x = lon, y = lat), fill = "red") +
  coord_cartesian(xlim = c(mean(biocro_results$lon) - 20, mean(biocro_results$lon) + 20), 
                  ylim = c(min(biocro_results$lat) - 5, max(biocro_results$lat) + 5)) +
  theme(legend.position = "none")
```

The final map plots estimated biomass values for the `r length(unique(biocro_results$latlon))` locations on this zoomed in map of east Asia, running through each day's values. 

```{r}
gif_background <- map_data("world") %>% 
  filter(region == "China" | region == "Russia")

colfunc <- colorRampPalette(c("yellow", "darkgreen"))
biomass_animation <- ggplot() +
  geom_polygon(data = gif_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = biocro_results, aes(x = lon, y = lat, fill = total_biomass)) +
    coord_cartesian(xlim = c(mean(biocro_results$lon) - 20, mean(biocro_results$lon) + 20), 
                  ylim = c(min(biocro_results$lat) - 5, max(biocro_results$lat) + 5)) +
  scale_fill_gradientn(colors = colfunc(10)) +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

animate(biomass_animation, fps = 100)
anim_save("biomass_animation_chiruss.gif", animation = biomass_animation, 
          path = "biocro_regional_darpa_files/", fps = 100)
```
