---
title: "How does duration of decreased stomatal slope affect transpiration and biomass?"
author: "Author: Jessica Guo"
editor_options: 
  chunk_output_type: console
---

# Overview
Recent data from the Gehan lab at Donald Danforth Plant Center demonstrated that in the psent-3 line, application of mandipropamid reduced leaf temperatures by an average of 0.18 °C for 48 hours. However, the Zhang lab demonstrated that Amax did not differ significantly between psent-3 and the wild type ME-034. Therefore, we conduct a sensitivity experiment on the duration of reduced stomatal conductance on Setaria biomass and transpiration, using the same set of physiological parameters as ME-034. 

Here, we wrote a function 'BioCro_restart()' which runs BioCro in 3 pieces: 1) original parameter set, 2) stomatal slope and cuticular conductance reduced by a proportion, and 3) original parameter set. The final biomass of one run is retained as the initial biomass of the subsequent run, yielding smooth accumulation of biomass across the experiment. 

```{r}
library(BioCro)
library(PEcAn.BIOCRO)
library(dplyr)
library(ggplot2)
library(cowplot)
source("Gs_sensitivity_inputs/BioCro_restart.R")
```


First, we will use hourly weather data from 2 locations, St. Louis (MO) and Pueblo (CO) extracted and used previously in 'Gs_testing.Rmd' and 'regional_runs.Rmd'. BioCro will be run for 100 days, with reduced stomatal parameters by 50% starting on day 20 for a duration of 1 to 21 days. Start run on June 1, 2010. 
```{r}
loc.coords <- data.frame(site = c("St. Louis, MO", "Pueblo, CO"),
                         lat = c(38.625, 38.375), 
                         lon = c(-90.125, -104.625))

duration <- 1:21

output.list <- list()
for(loc in 1:nrow(loc.coords)) {
  # Add met data
  biocro_met_path <- paste0("/data/dbfiles/biocro_met/stlew-",
                            loc.coords$lat[loc], "-", loc.coords$lon[loc], ".2010.csv")
  met <- read.csv(biocro_met_path) %>%
    filter(year == 2010 & doy >= 152)
  
  # Add biocro xml
  config_path <- paste0("Gs_testing_inputs/config_wt.xml")
  config <- read.biocro.config(config_path)
  
  # List of output
  output <- list()
  for(i in 1:length(duration)) {
    single <- BioCro_restart(met = met,
                      config = config, 
                      duration = duration[i],
                      start_day = 20,
                      end_day = 100,
                      prop = 0.5)
    ind <- as.vector(which(lapply(single, length) == (24*100)))
    output[[i]] <- do.call(cbind.data.frame, single[ind]) %>%
      mutate(duration = duration[i]) %>%
      relocate(duration)
  }
  
  # Combine across durations
  output.list[[loc]] <- do.call(rbind.data.frame, output) %>%
    mutate(location = loc.coords$site[loc]) %>%
    relocate(location)
  
  # Add instance of no reduction in gs parameters
  control <- BioGro(met, day1 = 1, dayn = 100,
                    soilControl = l2n(config$pft$soilControl),
                    canopyControl = l2n(config$pft$canopyControl),
                    phenoControl = l2n(config$pft$phenoParms),
                    seneControl = l2n(config$pft$seneControl),
                    photoControl = l2n(config$pft$photoParms))
  temp <- do.call(cbind.data.frame, control[ind]) %>%
      mutate(duration = 0,
             location = loc.coords$site[loc]) %>%
      relocate(location, duration)
  output.list[[loc]] <- rbind.data.frame(temp, output.list[[loc]])
  
  print(paste0("Finished with ", loc.coords$site[loc]))
}
  
output.df <- do.call(rbind.data.frame, output.list) %>%
  mutate(total_biomass = Leaf + Stem + Root,
         transpiration = CanopyTrans) %>%
  group_by(location, duration, DayofYear) %>%
  summarize(dailyBiomass = mean(total_biomass),
            dailyT = sum(transpiration)) %>%
  mutate(Date = seq(as.Date("2010-06-01"), as.Date("2010-09-08"), by = "day"))
```

The output is now a single dataframe object, hour timeseries for 100 days across 2 locations and 21 durations, plus a control instance of no change in parameters. 

One way to visualize this is biomass at the end of 100 days as a proportion of the control. 
```{r}
control <- output.df %>%
  ungroup() %>%
  filter(duration == 0) %>%
  rename(controlBiomass = dailyBiomass,
         controlT = dailyT) %>%
  dplyr::select(location, DayofYear, controlBiomass, controlT)

biomass_comp <- left_join(output.df %>% filter(duration != 0), control,
                          by = c("location", "DayofYear")) %>%
  mutate(biomassProp = dailyBiomass / controlBiomass,
         tProp = dailyT / controlT)

ggplot() +
  geom_point(data = control, aes(x = DayofYear, y = controlT, color = "control")) +
  geom_point(data = output.df[output.df$duration == 1,], 
             aes(x = DayofYear, y = dailyT, color = "reduced")) +
  facet_wrap(~location)

ggplot() +
  geom_point(data = control, aes(x = DayofYear, y = controlBiomass, color = "control")) +
  geom_point(data = output.df[output.df$duration == 1,], 
             aes(x = DayofYear, y = dailyBiomass, color = "reduced")) +
  facet_wrap(~location)

fig_biomass <- ggplot(biomass_comp, aes(x = Date)) +
  geom_tile(aes(y = duration, fill = biomassProp)) +
  facet_wrap(~location) +
  scale_fill_gradient2(low = "goldenrod", mid = "lightgray", 
                       high = "forestgreen", midpoint = 1) +
  scale_y_continuous("Days reduced") +
  theme_cowplot() +
  theme(strip.background =element_blank(),
        legend.title = element_blank(),
        plot.title = element_text(size = 12),
        axis.title.x = element_blank()) +
  labs(title = "Biomass proportion")

fig_trans <- ggplot(biomass_comp, aes(x = Date)) +
  geom_tile(aes(y = duration, fill = tProp)) +
  facet_wrap(~location) +
  scale_fill_gradient2(low = "pink", mid = "lightgray", 
                       high = "blueviolet", midpoint = 1) +
  scale_y_continuous("Days reduced") +
  theme_cowplot() +
  theme(strip.background =element_blank(),
        legend.title = element_blank(),
        plot.title = element_text(size = 12),
        axis.title.x = element_blank()) +
  labs(title = "Transpiration proportion")

plot_grid(fig_biomass, fig_trans, ncol = 1)
```
