---
title: "How does duration of decreased stomatal slope affect transpiration and biomass?"
author: "Author: Jessica Guo"
---

# Overview
Recent data from the Gehan lab at Donald Danforth Plant Center demonstrated that in the psent-3 line, application of mandipropamid reduced leaf temperatures by an average of 0.18 °C for 48 hours. However, the Zhang lab demonstrated that Amax did not differ significantly between psent-3 and the wild type ME-034. Therefore, we conduct a sensitivity experiment on the duration of reduced stomatal conductance on Setaria biomass and transpiration, using the same set of physiological parameters as ME-034. 

Here, we wrote a function 'BioCro_restart()' which runs BioCro in 3 pieces: 1) original parameter set, 2) stomatal slope and cuticular conductance reduced by a proportion, and 3) original parameter set. The final biomass of one run is retained as the initial biomass of the subsequent run, yielding smooth accumulation of biomass across the experiment. 

```{r}
library(BioCro)
library(PEcAn.BIOCRO)
library(dplyr)
source("Gs_sensitivity_inputs/BioCro_restart.R")
```


First, we will use hourly weather data from 2 locations, St. Louis (MO) and Pueblo (CO) extracted and used previously in 'Gs_testing.Rmd' and 'regional_runs.Rmd'. 
```{r}
loc.coords <- data.frame(site = c("St. Louis, MO", "Pueblo, CO"),
                         lat = c(38.625, 38.375), 
                         lon = c(-90.125, -104.625))

duration <- 1:21

output.list <- list()
for(loc in 1:nrow(loc.coords)) {
  # Add met data
  biocro_met_path <- paste0("/data/dbfiles/biocro_met/stlew-",
                            loc.coords$lat[loc], "-", loc.coords$lon[loc], ".2010.csv")
  met <- read.csv(biocro_met_path) %>%
    filter(year == 2010)
  
  # Add biocro xml
  config_path <- paste0("Gs_testing_inputs/config_wt.xml")
  config <- read.biocro.config(config_path)
  
  # List of output
  output <- list()
  for(i in 1:length(duration)) {
    single <- BioCro_restart(met = met,
                      config = config, 
                      duration = duration[i],
                      start_day = 50,
                      end_day = 100,
                      prop = 0.5)
    ind <- as.vector(which(lapply(single, length) == (24*100)))
    output[[i]] <- as.data.frame(do.call(cbind, single[ind])) %>%
      mutate(duration = duration[i]) %>%
      relocate(duration)
  }
  
  # Combine across durations
  output.list[[loc]] <- do.call(rbind.data.frame, output) %>%
    mutate(location = loc.coords$site[loc]) %>%
    relocate(location)
  
  print(paste0("Finished with ", loc.coords$site[loc]))
}
  
output.df <- do.call(rbind.data.frame, output.list)
```

