---
title: "How to Do Regional Runs on DARPA Data Using BioCro Model"
author: "Author: Kristina Riemer"
output: github_document
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

## Read in libraries

```{r}
devtools::load_all('~/pecan/models/biocro')
library(PEcAn.BIOCRO)
library(ncdf4)
library(PEcAn.all)
library(dplyr)
library(purrr)
library(udunits2)
library(ggplot2)
library(gganimate)
```
Here we combine Setaria WildType (ME034) with plants that have modified traits. One is with decreased stomatal conductance represented by a reduced slope in the slope parameter for the Ball Berry model of stomatal conductance. The second is for a dwarf plant.

**Decreased Stomatal Slope Regional Runs**

This set of runs uses the same parameters as above for the high light treatment, except in `config.xml` the stomatal slope input value is changed from `r config$pft$photoParms$b1` to `r config_ss$pft$photoParms$b1`. 

**Changed Carbon Allocation Regional Runs**

This set of runs uses the same parameters as above for the high light treatment, except in `config.xml` 

# Estimate biomass & transpiration

The three genotypes are 1) high light chamber treatment, 2) decreased stomatal slope values, and 3) modified carbon allocation to represent the dwarf plants. The first genotype's config file here comes from [the high light treatment BioCro run](https://github.com/az-digitalag/model-vignettes/blob/master/BioCro/DARPA/high_light.Rmd), specifically the config file in the SA-median run folder. The second has the stomatal slope value changed from 4.86 to 3.36. The third has the values of all `kStem` halved, with the remaining half split evenly between the other plant parts by stage. 

We also read in a soil file, which is not necessary but is useful. These data are from [the harmonized world soil database](http://www.fao.org/soils-portal/soil-survey/soil-maps-and-databases/harmonized-world-soil-database-v12/en/). 

```{r}
# genotypes <- c("wt", "ss", "ca")
genotypes <- c("wt", "ss")

regions <- c("stlew", "chiruss")

soil_nc <- nc_open("/data/dbfiles/soil/hwsd.nc")

grids <- list()
outdir <- list()
for(r in regions){
  # Open met data
  met <- nc_open(paste0("/data/dbfiles/narr_subsets/", r, ".nc"))
  # Create lat/lon grid
  grids[[r]] <- expand.grid(lat = ncvar_get(met, "latitude"),
                            lon = ncvar_get(met, "longitude"))
  # Create output directory filepath
  outdir[[r]] <- paste0("/data/output/biocro_regional_runs/", r)
  if(!dir.exists(outdir[[r]])){
  dir.create(outdir[[r]], 
             recursive = TRUE, 
             showWarnings = FALSE)
  }
}
```

The BioCro model is then run for each location of interest, using the config file, soil data, and weather data. The locations are determined and indexed from the weather data subset, which are then used to read in the relevant weather csv for the location of interest. 

 The weather data comes from [the CRUNCEP](https://rda.ucar.edu/datasets/ds314.3/) global gridded dataset provided by NCAR. The original NetCDF file has been subsetted according to [the weather subset vignette](https://github.com/az-digitalag/model-vignettes/blob/master/BioCro/DARPA/regional_runs_inputs/subset_weather.Rmd) and converted to the correct csv format required by BioCro according to [the convert weather vignette](https://github.com/az-digitalag/model-vignettes/blob/master/BioCro/DARPA/regional_runs_inputs/convert_weather_stlew.Rmd). This data is specifically for an extended longitudinal range across North America that includes the city of St. Louis.  

The results from running BioCro for each location produces hourly, daily, and yearly estimates for biomass, transpiration, etc. for the time range we specified in the config file. We are using only the daily values, combining them for all locations. These are saved as csv files. 

```{r}
# Only run ONCE
inputs <- expand.grid(genotype = genotypes, region = regions)
for(i in 3:nrow(inputs)) {
  genotype <- inputs$genotype[i]
  region <- inputs$region[i]
  config_path <- paste0("regional_runs_inputs/config_", genotype, ".xml")
  config <- read.biocro.config(config_path)
  regional_run_results <- c()
  for(p in 1:nrow(grids[[region]])){ 
    biocro_met_path <- paste0("/data/dbfiles/biocro_met/", region, "/", region, "-",
                            grids[[region]]$lat[p], "-", grids[[region]]$lon[p])
    regional_run_result <- run.biocro(grids[[region]]$lat[p], grids[[region]]$lon[p],
                               metpath = biocro_met_path,
                               soil.nc = soil_nc,
                               config = config)
    regional_run_result_daily <- regional_run_result$daily
    regional_run_result_daily$lat <- grids[[region]]$lat[p]
    regional_run_result_daily$lon <- grids[[region]]$lon[p]
    regional_run_results <- rbind(regional_run_results, regional_run_result_daily)
    print(p)
  }
  results_path <- paste0(outdir[[region]], "/", genotype, ".csv")
  write.csv(regional_run_results, results_path)
}
```

The csv files are read in, combined, and cleaned/modified for later plotting. First, combine for biomass and transpiration time series for all genotypes and regions. 

```{r}
inputs <- expand.grid(genotype = genotypes, region = regions)
rr_ts <- c()
for(i in 1:nrow(inputs)){
  genotype <- inputs$genotype[i]
  region <- inputs$region[i]
  
  results_path <- paste0(outdir[[region]], "/", genotype, ".csv")
  rr_single <- read.csv(results_path) %>%
    mutate(genotype = genotype,
           region = region)

  rr_ts <- rbind(rr_ts, rr_single)
}

rr_ts <- rr_ts %>% 
  mutate(date = as.Date(doy, "2009-12-31"), 
         latlon = paste0(lat, lon), 
         total_biomass = ud.convert(as.numeric(Stem + Leaf + Root + Rhizome + Grain), 
                                    "Mg/ha", "kg/m2"), 
         total_biomass = ifelse(total_biomass < 0, 0, total_biomass), 
         transpiration = ud.convert(CanopyTrans, "Mg/ha/hr", "kg/m2/hr"), 
         evaporation = ud.convert(SoilEvaporation, "Mg/ha/hr", "kg/m2/hr"), 
         T_ET = transpiration/ (transpiration + evaporation), 
         genotype_formal = case_when(
           genotype == "wt" ~ "Wild type", 
           genotype == "ss" ~ "Reduced stomatal sensitivity"
         ), 
         genotype_formal = factor(genotype_formal, levels = c("Wild type", 
                                                              "Reduced stomatal sensitivity")))
```

Then, compare triggered genotype (in this case 'ss') as a proportion of the wild type for each location. 
```{r}
rr_rel <- data.frame()
for(r in regions) {
  wt <- rr_ts %>%
    filter(genotype == "wt" & region == r) %>%
    mutate(date = as.POSIXct(date)) %>% 
    select(region, date, doy, lat, lon, latlon, total_biomass, transpiration, genotype_formal)
  ss <- rr_ts %>%
    filter(genotype == "ss" & region == r) %>%
    mutate(date = as.POSIXct(date)) %>%
    select(region, date, doy, lat, lon, latlon, total_biomass, transpiration, genotype_formal)
  
  rr_relative <- wt %>%
    mutate(trans_ss = ss$transpiration,
           biomass_ss = ss$total_biomass) %>%
    mutate(prop_trans = trans_ss/transpiration,
           prop_biomass = biomass_ss/total_biomass)
  
  rr_rel <- rbind.data.frame(rr_rel, rr_relative)
}
```

# Plot results of biomass and transpiration

## Time series by genotype

The relative total bioamss (the sum of stem, leaf, grain, rhizome, and root biomasses) and relative transpration by region are visualized. 

```{r}
ggplot(rr_rel, aes(x = date, y = prop_biomass, group = latlon, color = latlon)) +
  geom_line() +
  facet_wrap(~region) +
  guides(color = FALSE)
  
ggplot(rr_rel, aes(x = date, y = prop_trans, group = latlon, color = latlon)) +
  geom_line() +
  facet_wrap(~region) +
  guides(color = FALSE)
```

### N. America

The final GIF of estimated biomass/transpiration values for the `r nrow(grids$stlew)` locations on a map of the United States, running through each day's values. 

```{r}
na_background <- map_data("state")
STL_coords <- data.frame(name = "St. Louis", lon = -90.199402, lat = 38.627003)

biomass_ts <- ggplot() +
  geom_polygon(data = na_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = rr_ts[rr_ts$region == "stlew",], aes(x = lon, y = lat, fill = total_biomass), 
              alpha = 0.9) +
  geom_point(data = STL_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "yellow", mid = "yellowgreen", high = "darkgreen", 
                       midpoint = 1, 
                       breaks = seq(-10, 10, 0.4), 
                       limits = c(0, 2)) +
  coord_cartesian(xlim = c(-113, -80), ylim = c(36, 40.5)) +
  facet_wrap(~genotype_formal, ncol = 1) +
  theme_classic(base_size = 20) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey")) +
  labs(x = "", y = "", fill = "") +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

# animate(biomass_ts, fps = 10, height = 300, width = 600)
anim_save("biomass_ts.gif", animation = biomass_ts, 
          path = outdir$stlew, fps = 10, height = 300, width = 600)

transpiration_ts <- ggplot() +
  geom_polygon(data = na_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = rr_ts[rr_ts$region == "stlew",], aes(x = lon, y = lat, fill = transpiration), 
              alpha = 0.9) +
  geom_point(data = STL_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "lightblue", mid = "blue", high = "navyblue", 
                       midpoint = 7.5, 
                       breaks = seq(0, 20, 5), 
                       limits = c(0, 15)) +
  coord_cartesian(xlim = c(-113, -80), ylim = c(36, 40.5)) +
  facet_wrap(~genotype_formal, ncol = 1) +
  theme_classic(base_size = 20) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey")) +
  labs(x = "", y = "", fill = "") +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

# animate(transpiration_ts, fps = 10, height = 300, width = 600)
anim_save("trans_ts.gif", animation = transpiration_ts, 
          path = outdir$stlew, fps = 10, height = 300, width = 600)
```

### Eurasia

The final GIF of estimated biomass/transpiration values for the `r nrow(grids$chiruss)` locations on a map of Eurasia, running through each day's values. 

```{r}
EA_background <- map_data("world") %>% 
  filter(region == "China" | region == "Russia" | region == "Kazakhstan" | region == "Mongolia")
URU_coords <- data.frame(name = "Urumuqi", lon = 87.6168, lat = 43.8256)

biomass_ts <- ggplot() +
  geom_polygon(data = EA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = rr_ts[rr_ts$region == "chiruss",], aes(x = lon, y = lat, fill = total_biomass), 
              alpha = 0.9) +
  geom_point(data = URU_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "yellow", mid = "yellowgreen", high = "darkgreen", 
                       midpoint = 1, 
                       breaks = seq(-10, 10, 0.4), 
                       limits = c(0, 2)) +
  coord_cartesian(xlim = c(31, 186), ylim = c(20, 80)) +
  facet_wrap(~genotype_formal, ncol = 1) +
  theme_classic(base_size = 20) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey")) +
  labs(x = "", y = "", fill = "") +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

# animate(biomass_ts, fps = 10, height = 600, width = 600)
anim_save("biomass_ts.gif", animation = biomass_ts, 
          path = outdir$chiruss, fps = 10, height = 600, width = 600)

transpiration_ts <- ggplot() +
  geom_polygon(data = EA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = rr_ts[rr_ts$region == "chiruss",], aes(x = lon, y = lat, fill = transpiration), 
              alpha = 0.9) +
  geom_point(data = URU_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "lightblue", mid = "blue", high = "navyblue", 
                       midpoint = 7.5, 
                       breaks = seq(0, 20, 5), 
                       limits = c(0, 15)) +
  coord_cartesian(xlim = c(31, 186), ylim = c(20, 80)) +
  facet_wrap(~genotype_formal, ncol = 1) +
  theme_classic(base_size = 20) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey")) +
  labs(x = "", y = "", fill = "") +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

# animate(transpiration_ts, fps = 10, height = 600, width = 600)
anim_save("trans_ts.gif", animation = transpiration_ts, 
          path = outdir$chiruss, fps = 10, height = 600, width = 600)

```

## Relative biomass and transpiration proportions

### N. America

Comparison of the relative change in biomass and transpiration per pixel and over time. 
```{r}
NA_background <- map_data("state")
STL_coords <- data.frame(name = "St. Louis", lon = -90.199402, lat = 38.627003)

trans_prop <- ggplot() +
  geom_polygon(data = NA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = rr_rel[rr_rel$region == "stlew",], aes(x = lon, y = lat, fill = prop_trans), 
              alpha = 0.9) +
  geom_point(data = STL_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "maroon", mid = "lightblue", high = "darkblue", 
                       midpoint = 1, 
                       breaks = c(0.3, 1, 1.2), 
                       limits = c(0.3, 1.2)) +
  coord_cartesian(xlim = c(-113, -80), ylim = c(36, 40.5)) +
  theme_classic(base_size = 12) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey")) +
  labs(x = "", y = "", fill = "") +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

# animate(trans_prop, fps = 10, height = 150, width = 600)
anim_save("trans_rel_prop.gif", animation = trans_prop, 
          path = outdir$stlew, fps = 10, height = 150, width = 600)

biomass_prop <- ggplot() +
  geom_polygon(data = NA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = rr_rel[rr_rel$region == "stlew",], aes(x = lon, y = lat, fill = prop_biomass), 
              alpha = 0.9) +
  geom_point(data = STL_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "maroon", mid = "yellowgreen", high = "forestgreen", 
                       midpoint = 1, 
                       breaks = c(0.2, 1, 1.2), 
                       limits = c(0.2, 1.2)) +
  coord_cartesian(xlim = c(-113, -80), ylim = c(36, 40.5)) +
  theme_classic(base_size = 12) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey")) +
  labs(x = "", y = "", fill = "") +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

# animate(biomass_prop, fps = 10, height = 150, width = 600)
anim_save("biomass_rel_prop.gif", animation = biomass_prop, 
          path = outdir$stlew, fps = 10, height = 150, width = 600)
```

### Eurasia

Comparison of the relative change in biomass and transpiration per pixel and over time. 
```{r}
EA_background <- map_data("world") %>% 
  filter(region == "China" | region == "Russia" | region == "Kazakhstan" | region == "Mongolia")
URU_coords <- data.frame(name = "Urumuqi", lon = 87.6168, lat = 43.8256)

trans_prop <- ggplot() +
  geom_polygon(data = EA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = rr_rel[rr_rel$region == "chiruss",], aes(x = lon, y = lat, fill = prop_trans), 
              alpha = 0.9) +
  geom_point(data = URU_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "maroon", mid = "lightblue", high = "navyblue", 
                       midpoint = 1, 
                       breaks = c(0.6, 1, 1.2), 
                       limits = c(0.6, 1.2)) +
  coord_cartesian(xlim = c(85, 99), ylim = c(25, 80)) +
  theme_classic(base_size = 12) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey")) +
  labs(x = "", y = "", fill = "") +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

# animate(trans_prop, fps = 10, height = 500, width = 250)
anim_save("trans_rel_prop.gif", animation = trans_prop, 
          path = outdir$chiruss, fps = 10, height = 500, width = 250)

biomass_prop <- ggplot() +
  geom_polygon(data = EA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = rr_rel[rr_rel$region == "chiruss",], aes(x = lon, y = lat, fill = prop_biomass), 
              alpha = 0.9) +
  geom_point(data = URU_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "maroon", mid = "yellowgreen", high = "darkgreen", 
                       midpoint = 1, 
                       breaks = c(0.6, 1, 1.2), 
                       limits = c(0.6, 1.2)) +
  coord_cartesian(xlim = c(80, 99), ylim = c(25, 80)) +
  theme_classic(base_size = 12) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey")) +
  labs(x = "", y = "", fill = "") +
  transition_manual(doy) +
  ggtitle('Day: {current_frame}')

# animate(biomass_prop, fps = 10, height = 300, width = 150)
anim_save("biomass_rel_prop.gif", animation = biomass_prop, 
          path = outdir$chiruss, fps = 10, height = 500, width = 250)
```

## Static time series of relative differences

### N. America
```{r}
NA_static <- rr_rel[rr_rel$region == "stlew",] %>%
  mutate(mon = format(as.Date(date), "%m"), 
         day = format(as.Date(date), "%d")) %>%
  filter(day == "01")

NA_background <- map_data("state")
STL_coords <- data.frame(name = "St. Louis", lon = -90.199402, lat = 38.627003)

trans_grid <- ggplot() +
  geom_polygon(data = NA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = NA_static[NA_static$region == "stlew",], aes(x = lon, y = lat, fill = prop_trans), 
              alpha = 0.9) +
  geom_point(data = STL_coords, aes(x = lon, y = lat), size = 0.5) +
  scale_fill_gradient2(low = "maroon", mid = "lightblue", high = "darkblue", 
                       midpoint = 1, 
                       breaks = c(0.3, 1, 1.2), 
                       limits = c(0.3, 1.2)) +
  coord_cartesian(xlim = c(-113, -80), ylim = c(36, 40.5)) +
  scale_y_continuous(breaks = seq(36,40, 2)) +
  facet_wrap(~mon, ncol = 3) +
  theme_bw(base_size = 12) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey"),
        strip.background = element_blank()) +
  labs(x = "", y = "", fill = "")
jpeg(filename = paste0(outdir$stlew, "/trans_rel_grid.jpeg"), 
     height = 3.5, width = 8,  units = "in", res = 600)
print(trans_grid)
dev.off()

biomass_grid <- ggplot() +
  geom_polygon(data = NA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = NA_static[NA_static$region == "stlew",], aes(x = lon, y = lat, fill = prop_biomass), 
              alpha = 0.9) +
  geom_point(data = STL_coords, aes(x = lon, y = lat), size = 0.5) +
  scale_fill_gradient2(low = "maroon", mid = "yellowgreen", high = "forestgreen", 
                       midpoint = 1, 
                       breaks = c(0.2, 1, 1.2), 
                       limits = c(0.2, 1.2)) +
  coord_cartesian(xlim = c(-113, -80), ylim = c(36, 40.5)) +
  scale_y_continuous(breaks = seq(36,40, 2)) +
  facet_wrap(~mon, ncol = 3) +
  theme_bw(base_size = 12) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey"),
        strip.background = element_blank()) +
  labs(x = "", y = "", fill = "")
jpeg(filename = paste0(outdir$stlew, "/biomass_rel_grid.jpeg"), 
     height = 3.5, width = 8,  units = "in", res = 600)
print(biomass_grid)
dev.off()
```

### Eurasia
```{r}
EA_static <- rr_rel[rr_rel$region == "chiruss",] %>%
  mutate(mon = format(as.Date(date), "%m"), 
         day = format(as.Date(date), "%d")) %>%
  filter(day == "01")

EA_background <- map_data("world") %>% 
  filter(region == "China" | region == "Russia" | region == "Kazakhstan" | region == "Mongolia")
URU_coords <- data.frame(name = "Urumuqi", lon = 87.6168, lat = 43.8256)

trans_grid <- ggplot() +
  geom_polygon(data = EA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = EA_static[EA_static$region == "chiruss",], aes(x = lon, y = lat, fill = prop_trans), 
              alpha = 0.9) +
  geom_point(data = URU_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "maroon", mid = "lightblue", high = "navyblue", 
                       midpoint = 1, 
                       breaks = c(0.6, 1, 1.2), 
                       limits = c(0.6, 1.2)) +
  coord_cartesian(xlim = c(85, 99), ylim = c(25, 80)) +
  facet_wrap(~mon, ncol = 6) +
  theme_bw(base_size = 12) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey"),
        strip.background = element_blank()) +
  labs(x = "", y = "", fill = "") 

jpeg(filename = paste0(outdir$chiruss, "/trans_rel_grid.jpeg"), 
     height = 6, width = 6,  units = "in", res = 600)
print(trans_grid)
dev.off()

biomass_grid <- ggplot() +
  geom_polygon(data = EA_background, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_raster(data = EA_static[EA_static$region == "chiruss",], aes(x = lon, y = lat, fill = prop_biomass), 
              alpha = 0.9) +
  geom_point(data = URU_coords, aes(x = lon, y = lat)) +
  scale_fill_gradient2(low = "maroon", mid = "yellowgreen", high = "darkgreen", 
                       midpoint = 1, 
                       breaks = c(0.6, 1, 1.2), 
                       limits = c(0.6, 1.2)) +
  coord_cartesian(xlim = c(85, 99), ylim = c(25, 80)) +
  facet_wrap(~mon, ncol = 6) +
  theme_bw(base_size = 12) +
  theme(panel.background = element_rect(fill = "grey", colour = "grey"), 
        panel.grid.major = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "grey"),
        strip.background = element_blank()) +
  labs(x = "", y = "", fill = "") 

jpeg(filename = paste0(outdir$chiruss, "/biomass_rel_grid.jpeg"), 
     height = 6, width = 6,  units = "in", res = 600)
print(biomass_grid)
dev.off()
```
