---
title: "Community Dynamics Simulation"
author: "Kristina Riemer"
date: "3/25/2019"
output:
  pdf_document: default
  html_document: default
---

## Run model

The Ecosystem Demography ([ED2](https://github.com/EDmodel/ED2)) model was run using the [PEcAn VM in Virtualbox](https://pecanproject.github.io/pecan-documentation/develop/GettingStarted.html#installing-and-running-pecan). The following settings were used: 

 * Host = pecan.vm
 * Model = ED2.2 (git)
 * Site Group = All Sites
 * Site = EBI Energy farm
 * PFTs
    1. ebifarm.c3grass
    2. ebifarm.c4crop
    3. ebifarm.forb
 * Start Date = 2004/01/01
 * End Date = 2009/12/31
 * ED2.cohort = ebifarm.lat40.0long-88.0.css
 * ED2.patch = ebifarm.lat40lon-88.0.pss
 * ED2.site = ebifarm. lat40.0lon88.0.site
 * Ed.met_driver_header = ED_MET_DRIVER_HEADER 2004-2009
 * Land use = Earth Land Surface
 * Soil = FAO_
 * Thermal sums = Earth Land Surface
 * Vegetation = OGE2

The configuration file ED2IN was also edited by checking the "Edit config file" box. The following three sections were changed. 

```{sh, eval=FALSE}
   !---------------------------------------------------------------------------------------!
   ! ED2 File output.  For all the variables 0 means no output and 3 means HDF5 output.    !
   !                                                                                       !
   ! IFOUTPUT -- Fast analysis.  These are mostly polygon-level averages, and the time     !
   !             interval between files is determined by FRQANL                            !
   ! IDOUTPUT -- Daily means (one file per day)                                            !
   ! IMOUTPUT -- Monthly means (one file per month)                                        !
   ! IQOUTPUT -- Monthly means of the diurnal cycle (one file per month).  The number      !
   !             of points for the diurnal cycle is 86400 / FRQANL                         !
   ! IYOUTPUT -- Annual output.                                                            !
   ! ITOUTPUT -- Instantaneous fluxes, mostly polygon-level variables, one file per year.  !
   ! IOOUTPUT -- Observation time output. Equivalent to IFOUTPUT, except only at the       !
   !             times specified in OBSTIME_DB.                                            !
   ! ISOUTPUT -- restart file, for HISTORY runs.  The time interval between files is       !
   !             determined by FRQHIS                                                      !
   !---------------------------------------------------------------------------------------!
   NL%IFOUTPUT = 0
   NL%IDOUTPUT = 0
   NL%IMOUTPUT = 3
   NL%IQOUTPUT = 0
   NL%IYOUTPUT = 0
   NL%ITOUTPUT = 0
   NL%IOOUTPUT = 0
   NL%ISOUTPUT = 0
   !---------------------------------------------------------------------------------------!
```

```{sh, eval=FALSE}
   !---------------------------------------------------------------------------------------!
   ! ATTACH_METADATA -- Flag for attaching metadata to HDF datasets.  Attaching metadata   !
   !                    will aid new users in quickly identifying dataset descriptions but !
   !                    will compromise I/O performance significantly.                     !
   !                    0 = no metadata, 1 = attach metadata                               !
   !---------------------------------------------------------------------------------------!
   NL%ATTACH_METADATA = 1
   !---------------------------------------------------------------------------------------!
```

```{sh, eval=FALSE}
   !---------------------------------------------------------------------------------------!
   ! UNITFAST  --  The following variables control the units for FRQFAST/OUTFAST, and      !
   ! UNITSTATE     FRQSTATE/OUTSTATE, respectively.  Possible values are:                  !
   !               0.  Seconds;                                                            !
   !               1.  Days;                                                               !
   !               2.  Calendar months (variable)                                          !
   !               3.  Calendar years  (variable)                                          !
   !                                                                                       !
   ! N.B.: 1. In case OUTFAST/OUTSTATE are set to special flags (-1 or -2)                 !
   !          UNITFAST/UNITSTATE will be ignored for them.                                 !
   !       2. In case IQOUTPUT is set to 3, then UNITFAST has to be 0.                     !
   !                                                                                       !
   !---------------------------------------------------------------------------------------!
   NL%UNITFAST = 1
   NL%UNITSTATE = 1
   !---------------------------------------------------------------------------------------!

   !---------------------------------------------------------------------------------------!
   ! OUTFAST/OUTSTATE -- these control the number of times per file.                       !
   !                      0. Each time gets its own file                                   !
   !                     -1. One file per day                                              !
   !                     -2. One file per month                                            !
   !                    > 0. Multiple timepoints can be recorded to a single file reducing !
   !                         the number of files and i/o time in post-processing.          !
   !                         Multiple timepoints should not be used in the history files   !
   !                         if you intend to use these for HISTORY runs.                  !
   !---------------------------------------------------------------------------------------!
   NL%OUTFAST = 0
   NL%OUTSTATE = 0
   !---------------------------------------------------------------------------------------!
```

## Visualize results

All the Output files of the format `analysis-E-*-*-00-000000-g01.h5` were downloaded by selecting them from the dropdown menu and selecting the "Show Output File" button. 

There is one file per month. The data format for a single month for the variable `MMEAN_LAI_PY` is shown below. Each row is a PFT and each column is a cohort. 

```{r, message=FALSE}
library(ncdf4)
single_nc <- nc_open("ed2_results/analysis-E-2004-01-00-000000-g01.h5")
single_nc_lai <- ncvar_get(single_nc, "MMEAN_LAI_PY")
single_nc_lai
```

Each monthly LAI dataset is summed across cohort and only the rows for the three PFTs of interest are retained. All the monthly datasets are combined into one. 

```{r, message=FALSE}
library(dplyr)
library(tibble)

extract_lai <- function(file_path){
  nc <- nc_open(file_path)
  lai <- ncvar_get(nc, "MMEAN_LAI_PY")
  lai_df <- data.frame(lai)
  clean_lai_df <- lai_df %>% 
    transmute(LAI = rowMeans(.)) %>% 
    slice(1:3) %>% 
    add_column(PFT = c("c3grass", "c4crop", "forb"), 
             date = substr(file_path, 24, 30)) %>% 
    mutate(date = as.Date(paste0(date, "-01")))
}

nc_files <- list.files("ed2_results/", pattern = "*.h5")
nc_file_paths <- paste0("ed2_results/", nc_files)

all_lai <- lapply(nc_file_paths, extract_lai)
all_lai <- do.call(rbind, all_lai)
head(all_lai)
```

The LAI time series per PFT is plotted. 

```{r, message=FALSE}
library(ggplot2)

ggplot(all_lai, aes(x = date, y = LAI, color = PFT)) +
  geom_line()
```
