---
title: "CLM Sensitivity Analysis"
output: github_document
---

```{r}
library(broom)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

Data are parameter values with corresponding response variable values. 

```{r}
vegc <- readr::read_csv("clm5_CIG_C4grass_e101_Annual_mean_TOTVEGC_for_decomposition.csv")
```

### Elasticity

Sensitivity is the change in the response variable given a change in an input variable, which can be a parameter, initial condition, etc. This is representated by the derivative dY/dX. This is calculated below from the slope of the linear regression between the parameter values and response variable. 

```{r}
slopes <- tidy(lm(data = vegc, formula = Annual_mean_TOTVEGC ~ .)) %>% 
  rename(sensitivity = estimate)
```

Because the units of the derviative are determined by X's units, elasticity is used to standardize sensitivity across variables. Elasticity is dY/dX * (mean X/mean Y). 

```{r}
mean_vegc_response <- mean(vegc$Annual_mean_TOTVEGC)

mean_vegc_params <- vegc %>% 
  select(-Annual_mean_TOTVEGC) %>% 
  summarise_all(funs(mean)) %>% 
  gather(parameter, mean) %>% 
  mutate(elasticity_multiplier = mean / mean_vegc_response)

elasticity <- left_join(slopes, mean_vegc_params, by = c("term" = "parameter")) %>% 
  mutate(elasticity = sensitivity * elasticity_multiplier)
```

Elasticity values increasingly farther from zero mean change in x is causing increasingly greater change in y. A change in x causes the same change in y when elasticity is one. 

```{r}
ggplot(elasticity, aes(x = elasticity, y = term)) +
  geom_point() +
  geom_vline(xintercept = 0) +
  lims(x = c(-1, 1))
```

### Coefficient of Variation

CV is the normalized parameter variance of all the input parameters. 

```{r}
CVs <- vegc %>% 
  select(-Annual_mean_TOTVEGC) %>% 
  gather(parameter, value) %>% 
  group_by(parameter) %>% 
  summarize(mean = mean(value), 
            var = var(value)) %>% 
  mutate(sd = sqrt(var), 
         cv = (sd / mean) * 100)
```

```{r}
ggplot(CVs, aes(x = cv, y = parameter)) +
  geom_point() +
  geom_vline(xintercept = 0)
```

### Explained Standard Deviation

How much of the response variable uncertainty is explained by each of the parameters. Incorporates both parameter variance and sensitivity. 

```{r}
SDs <- tidy(aov(Annual_mean_TOTVEGC ~ ., data = vegc)) %>% 
  mutate(sd = sqrt(sumsq)) %>% 
  select(term, sd)
```

```{r}
ggplot(SDs, aes(x = sd, y = term)) +
  geom_point() +
  geom_vline(xintercept = 0)
```
