---
title: "Light and CO2 responses of gas exchange and NPQ"
author: "Jessica Guo"
date: "12/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Two sets of figures for the manuscript comparing regular and high night temperature for Setaria. The first figure is a set of standard light and CO2 response curves, while the second shows a selection of NPQ parameters in response to light. 

```{r library}
library(dplyr)
library(data.table)
library(ggplot2)
library(cowplot)
library(wesanderson)
experiments <- read.csv("../../sentinel-detection/data/cleaned_data/experiments.csv")
```

## Light and CO2 responses

Obtain ACi and AQ data from the sentinel-detection [repo](https://github.com/danforthcenter/sentinel-detection) for the relevant select treatments and combine with appropriate labels. 

```{r}
# Read in high night temp
hn_id <- experiments$ID[which(experiments$Treatment == "31_31_250")]
hn_aci <-  vector("list", length = length(hn_id))
hn_aq <-  vector("list", length = length(hn_id))
for(i in 1:length(hn_id)){
  aci <- read.csv(paste0("../../sentinel-detection/data/cleaned_data/ACi/A_Ci_curves_", hn_id[i], ".csv"), 
                  na.strings = "#DIV/0!")
  hn_aci[[i]] <-cbind.data.frame(ID = rep(hn_id[i], nrow(aci)), aci)
  aq <- read.csv(paste0("../../sentinel-detection/data/cleaned_data/AQ/AQin_curves_", hn_id[i], ".csv"), 
                  na.strings = "#DIV/0!")
  hn_aq[[i]] <-cbind.data.frame(ID = rep(hn_id[i], nrow(aq)), aq)
}

# Read in regular night temp
rn_id <- experiments$ID[which(experiments$Treatment == "31_22_250")]
rn_aci <-  vector("list", length = length(rn_id))
rn_aq <-  vector("list", length = length(rn_id))
for(i in 1:length(rn_id)){
  aci <- read.csv(paste0("../../sentinel-detection/data/cleaned_data/ACi/A_Ci_curves_", rn_id[i], ".csv"), 
                  na.strings = "#DIV/0!")
  rn_aci[[i]] <-cbind.data.frame(ID = rep(rn_id[i], nrow(aci)), aci)
  aq <- read.csv(paste0("../../sentinel-detection/data/cleaned_data/AQ/AQin_curves_", rn_id[i], ".csv"), 
                  na.strings = "#DIV/0!")
  rn_aq[[i]] <-cbind.data.frame(ID = rep(rn_id[i], nrow(aq)), aq)
}

# Only preserve identical column names
cnames <- Reduce(intersect, c(lapply(hn_aci, colnames), 
                                   lapply(rn_aci, colnames),
                                   lapply(hn_aq, colnames), 
                                   lapply(rn_aq, colnames)))
keep_match <- function(x) {x[, colnames(x) %in% cnames]}
hn_aci <- lapply(hn_aci, keep_match)
rn_aci <- lapply(rn_aci, keep_match)
hn_aq <- lapply(hn_aq, keep_match)
rn_aq <- lapply(rn_aq, keep_match)

# Bind with treatment identifier
aci_df <- rbind.data.frame(data.frame(Treatment = "31_31_250", do.call(rbind, hn_aci)),
                           data.frame(Treatment = "31_22_250", do.call(rbind, rn_aci)))
aq_df <- rbind.data.frame(data.frame(Treatment = "31_31_250", do.call(rbind, hn_aq)),
                           data.frame(Treatment = "31_22_250", do.call(rbind, rn_aq)))
# Keep relevant columns
aci <- aci_df %>%
  select(Treatment, ID, species, rep, obs, date, E, A, Ca, Ci, Pci, Pca) %>%
  mutate(trt = case_when(Treatment == "31_22_250" ~ "Control",
                         Treatment == "31_31_250" ~ "High night T"))

aq <- aq_df %>%
  select(Treatment, ID, species, rep, obs, date, E, A, Ca, Ci, Pci, Pca, Qin) %>%
  mutate(trt = case_when(Treatment == "31_22_250" ~ "Control",
                         Treatment == "31_31_250" ~ "High night T"))
```

To plot as Aci and AQ curves, summarize across replicates and for each Ci or Qin level. 

```{r}
# Set CO2 chamber conditions, match to actual Ca
clevels <- c(400, 300, 200, 100, 50, 20, 10, 
             500, 600, 700, 800, 900, 1000, 1200, 1500)
nearest_clevel <- function(x) {clevels[which.min(abs(x - clevels))]}
aci$roundCa <- c()
for(i in 1:nrow(aci)) {
  aci$roundCa[i] <- nearest_clevel(aci$Ca[i])
}

# set Qin chamber conditions, match to actual Qin
qlevels <- c(0, 50, 100, 200, 400, 
             600, 800, 1200, 1500)
nearest_qlevel <- function(x) {qlevels[which.min(abs(x - qlevels))]}
aq$roundQin <- c()
for(i in 1:nrow(aq)) {
  aq$roundQin[i] <- nearest_qlevel(aq$Qin[i])
}

# Functions to summarize
SE <- function(x) {sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x)))}
num <- function(x) {sum(!is.na(x))}

# Summarize Ci and A by roundCa
aci <- as.data.table(aci)
aci_sum <- aci[, .(Ci.m = mean(Ci,  na.rm = TRUE),
                   Ci.se = SE(Ci),
                   A.m = mean(A, na.rm = TRUE),
                   A.se = SE(A),
                   n = num(A)), by = .(trt, roundCa)]

# Summarize Qin and A by roundQin
aq <- as.data.table(aq)
aq_sum <- aq[, .(Qin.m = mean(Qin,  na.rm = TRUE),
                 Qin.se = SE(Qin),
                 A.m = mean(A, na.rm = TRUE),
                 A.se = SE(A),
                 n = num(A)), by = .(trt, roundQin)]

# Plot
cols <- wes_palette("GrandBudapest1")
fig1a <- ggplot(aci_sum, aes(x = Ci.m, y = A.m, color = trt)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = A.m - A.se, ymax = A.m + A.se)) +
  geom_errorbarh(aes(xmin = Ci.m - Ci.se, xmax = Ci.m + Ci.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste(C[i], " (", mu, "mol ", mol^-1, ")"))) +
  scale_y_continuous(expression(paste(A[n], " (", mu, "mol ", m^-2, " ", s^-1, ")"))) +
  scale_color_manual(values = cols) +
  theme(legend.position = c(0.7, 0.2),
        legend.title = element_blank())

fig1b <- ggplot(aq_sum, aes(x = Qin.m, y = A.m, color = trt)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = A.m - A.se, ymax = A.m + A.se)) +
  geom_errorbarh(aes(xmin = Qin.m - Qin.se, xmax = Qin.m + Qin.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste("PPFD (", mu, "mol ",  m^-2, " ", s^-1, ")"))) +
  scale_y_continuous(expression(paste(A[n], " (", mu, "mol ", m^-2, " ", s^-1, ")"))) +
  scale_color_manual(values = cols) +
  guides(color = FALSE)

jpeg(filename = "plots/Fig1_gas_ex_curves.jpg", height = 3, width = 8, units = "in", res = 600)
plot_grid(fig1a, fig1b, labels = c('A', 'B'), label_size = 12)
dev.off()
```


## NPQ plots

From the aq_df, extract relevant NPQ parameters and perform calculations for the C4 pathway. Adaptation of Tonantzin's original processing code. 

```{r}
# Include additional columns related to NPQ
npq <- aq_df %>%
  select(Treatment, ID, species, rep, obs, date, TIME,
         E, A, Ca, Ci, Pci, Pca, Qin,
         Fo, Fs, Fm, Fm., Fo., PhiPS2, Qabs_fs, Fv., PhiCO2,
         qP, qN) %>%
  mutate(trt = case_when(Treatment == "31_22_250" ~ "Control",
                         Treatment == "31_31_250" ~ "High night T"))

# Split by ID and rep (individual plant) to perform corrections
npq$entity <- paste0(npq$ID, "_", npq$rep)
npq_list <- split(npq, f = npq$entity)
for(i in 1:length(npq_list)) {
  # Corrected values
  npq_list[[i]] <- npq_list[[i]] %>%
    mutate(Fs_c = Fs - Fo/2, # Fs steady state fluorescence,  Fo in dark-adapted leaves
           Fm._c = Fm. - Fo/2, # Fm' (Fm.) maximal fluorescence from light-adapted leaves
           Fo._c = Fo. - Fo/2, # Minimum fluorescence intensity in the light-acclimated state
           PhiPS2_c = 1 - (Fs_c /Fm._c),
           ETR_c = PhiPS2_c * 0.5 * Qabs_fs,
           NPQ_c = (Fm - Fm._c) / Fm._c, # Non-photochemical quenching
           Fv._Fm. = Fv. / Fm._c,
           PhiPS2max = 1 - (Fo._c / Fm._c), # Maximum efficiency of PSII in light if all centers were open
           seconds = TIME - first(TIME)) # Seconds elapsed since first measurement
  
  #Remove outliers
  npq_list[[i]] $NPQ_c[which(npq_list[[i]] $NPQ_c < 0)] = NA
  npq_list[[i]] $PhiPS2max[which(npq_list[[i]] $PhiPS2max  < 0.1)] = NA
  npq_list[[i]] $PhiCO2[which(npq_list[[i]] $PhiCO2  < 0)] = NA
}

npq <- as.data.table(do.call(rbind, npq_list))
```


To plot as Aci and AQ curves, summarize across replicates and for each Ci or Qin level. 
```{r}
# Round light levels
for(i in 1:nrow(npq)) {
  npq$roundQin[i] <- nearest_qlevel(npq$Qin[i])
}

# Summarize
npq_sum <- npq[roundQin !=0 , .(A.m = mean(A, na.rm = TRUE), A.se = SE(A),
                   Qin.m = mean(Qin,  na.rm = TRUE), Qin.se = SE(Qin),
                   NPQ.m = mean(NPQ_c, na.rm = TRUE), NPQ.se = SE(NPQ_c),
                   ETR.m = mean(ETR_c, na.rm = TRUE), ETR.se = SE(ETR_c),
                   PhiPS2.m = mean(PhiPS2_c, na.rm = TRUE), PhiPS2.se = SE(PhiPS2_c),
                   PhiPS2max.m = mean(PhiPS2max, na.rm = TRUE), PhiPS2max.se = SE(PhiPS2max),
                   qP.m =  mean(qP, na.rm = TRUE), qP.se = SE(qP),
                   qN.m = mean(qN, na.rm = TRUE), qN.se = SE(qN),
                   FvFm.m = mean(Fv._Fm., na.rm = TRUE), FvFm.se = SE(Fv._Fm.)),  by = .(trt, roundQin)]

# Panel of 6 plots
fig2a <- ggplot(npq_sum, aes(x = Qin.m, y = NPQ.m, color = trt, group = trt)) +
  geom_point(size = 2) +
  geom_line(lty = 2) +
  geom_errorbar(aes(ymin = NPQ.m - NPQ.se, ymax = NPQ.m + NPQ.se)) +
  geom_errorbarh(aes(xmin = Qin.m - Qin.se, xmax = Qin.m + Qin.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste("PPFD (", mu, "mol ",  m^-2, " ", s^-1, ")"))) +
  scale_y_continuous(expression(paste("NPQ"))) +
  scale_color_manual(values = cols)  +
  theme(legend.position = c(0.65, 0.2),
        legend.title = element_blank())
fig2b <- ggplot(npq_sum, aes(x = Qin.m, y = ETR.m, color = trt, group = trt)) +
  geom_point(size = 2) +
  geom_line(lty = 2) +
  geom_errorbar(aes(ymin = ETR.m - ETR.se, ymax = ETR.m + ETR.se)) +
  geom_errorbarh(aes(xmin = Qin.m - Qin.se, xmax = Qin.m + Qin.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste("PPFD (", mu, "mol ",  m^-2, " ", s^-1, ")"))) +
  scale_y_continuous(expression(paste("ETR"))) +
  scale_color_manual(values = cols)  +
  guides(color = FALSE)
fig2c <- ggplot(npq_sum, aes(x = Qin.m, y = PhiPS2.m, color = trt, group = trt)) +
  geom_point(size = 2) +
  geom_line(lty = 2) +
  geom_errorbar(aes(ymin = PhiPS2.m - PhiPS2.se, ymax = PhiPS2.m + PhiPS2.se)) +
  geom_errorbarh(aes(xmin = Qin.m - Qin.se, xmax = Qin.m + Qin.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste("PPFD (", mu, "mol ",  m^-2, " ", s^-1, ")"))) +
  scale_y_continuous(expression(paste(Phi[PSII]))) +
  scale_color_manual(values = cols)  +
  guides(color = FALSE)
fig2d <- ggplot(npq_sum, aes(x = Qin.m, y = PhiPS2max.m, color = trt, group = trt)) +
  geom_point(size = 2) +
  geom_line(lty = 2) +
  geom_errorbar(aes(ymin = PhiPS2max.m - PhiPS2max.se, ymax = PhiPS2max.m + PhiPS2max.se)) +
  geom_errorbarh(aes(xmin = Qin.m - Qin.se, xmax = Qin.m + Qin.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste("PPFD (", mu, "mol ",  m^-2, " ", s^-1, ")"))) +
  scale_y_continuous(expression(paste(Phi[PSII[max]]))) +
  scale_color_manual(values = cols)  +
  guides(color = FALSE)
fig2e <- ggplot(npq_sum, aes(x = Qin.m, y = qP.m, color = trt, group = trt)) +
  geom_point(size = 2) +
  geom_line(lty = 2) +
  geom_errorbar(aes(ymin = qP.m - qP.se, ymax = qP.m + qP.se)) +
  geom_errorbarh(aes(xmin = Qin.m - Qin.se, xmax = Qin.m + Qin.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste("PPFD (", mu, "mol ",  m^-2, " ", s^-1, ")"))) +
  scale_y_continuous(expression(paste("qP"))) +
  scale_color_manual(values = cols)  +
  guides(color = FALSE)
fig2f <- ggplot(npq_sum, aes(x = Qin.m, y = qN.m, color = trt, group = trt)) +
  geom_point(size = 2) +
  geom_line(lty = 2) +
  geom_errorbar(aes(ymin = qN.m - qN.se, ymax = qN.m + qN.se)) +
  geom_errorbarh(aes(xmin = Qin.m - Qin.se, xmax = Qin.m + Qin.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste("PPFD (", mu, "mol ",  m^-2, " ", s^-1, ")"))) +
  scale_y_continuous(expression(paste("qN"))) +
  scale_color_manual(values = cols)  +
  guides(color = FALSE)

jpeg(filename = "plots/Fig2_NPQ.jpg", height = 8, width = 6, units = "in", res = 600)
plot_grid(fig2a, fig2b, fig2c, fig2d, fig2e, fig2f, ncol = 2, labels = LETTERS[1:6], label_size = 12)
dev.off()

# Bonus plot of Fv/Fm
fig2e <- ggplot(npq_sum, aes(x = FvFm.m, y = qN.m, color = trt, group = trt)) +
  geom_point(size = 2) +
  geom_line(lty = 2) +
  geom_errorbar(aes(ymin = qN.m - qN.se, ymax = qN.m + qN.se)) +
  geom_errorbarh(aes(xmin = FvFm.m - FvFm.se, xmax = FvFm.m + FvFm.se)) +
  theme_cowplot(11) +
  scale_x_continuous(expression(paste("PPFD (", mu, "mol ",  m^-2, " ", s^-1, ")"))) +
  scale_y_continuous(expression(paste("Fv/Fm"))) +
  scale_color_manual(values = cols)  +
  guides(color = FALSE)
```


