---
title: "Light and CO2 responses of gas exchange and NPQ"
author: "Jessica Guo"
date: "12/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This code produces parameter tables with separate columns for control and high night temperature. Parameter means and SE will be reported in a summary table for photosynthetic and stomatal parameters

```{r library}
library(dplyr)
library(data.table)
library(tidyr)
library(gt)
library(Hotelling)
```

## Summarize parameter data

Obtain ACi and AQ data from the sentinel-detection [repo](https://github.com/danforthcenter/sentinel-detection) and calculate A and gs at average conditions. In this case, combine ACi and AQ data into single dataframe and seek data from standard CO2 and light conditions of 400 ppm and PAR = 1500. 

```{r}
# List of experiments
experiments <- read.csv("~/sentinel-detection/data/cleaned_data/experiments.csv")

#
# Read in high night temp
hn_id <- experiments$ID[which(experiments$Treatment == "31_31_250")]
hn <-  vector("list", length = length(hn_id))
for(i in 1:length(hn_id)){
  aci <- read.csv(paste0("~/sentinel-detection/data/cleaned_data/ACi/A_Ci_curves_", hn_id[i], ".csv"), 
                  na.strings = "#DIV/0!")
  aq <- read.csv(paste0("~/sentinel-detection/data/cleaned_data/AQ/AQin_curves_", hn_id[i], ".csv"), 
                  na.strings = "#DIV/0!")
  
  dat <- rbind.data.frame(aci[, intersect(colnames(aci), colnames(aq))], 
                          aq[, intersect(colnames(aci), colnames(aq))])
  dat$ID = rep(hn_id[i], nrow(dat))
  dat$Treatment <- rep("31_31_250", nrow(dat))
  dat$trt <- rep("High night T", nrow(dat))
  hn[[i]] <- dat
}

# Read in regular night temp
rn_id <- experiments$ID[which(experiments$Treatment == "31_22_250")]
rn <-  vector("list", length = length(rn_id))
for(i in 1:length(rn_id)){
  aci <- read.csv(paste0("~/sentinel-detection/data/cleaned_data/ACi/A_Ci_curves_", rn_id[i], ".csv"), 
                  na.strings = "#DIV/0!")
  aq <- read.csv(paste0("~/sentinel-detection/data/cleaned_data/AQ/AQin_curves_", rn_id[i], ".csv"), 
                  na.strings = "#DIV/0!")
  
  dat <- rbind.data.frame(aci[, intersect(colnames(aci), colnames(aq))], 
                          aq[, intersect(colnames(aci), colnames(aq))])
  dat$ID <- rep(rn_id[i], nrow(dat))
  dat$Treatment <- rep("31_22_250", nrow(dat))
  dat$trt <- rep("Control", nrow(dat))
  rn[[i]] <- dat
}

# Column names still uneven
cnames <- Reduce(intersect, c(lapply(hn, colnames), 
                              lapply(rn, colnames)))
keep_match <- function(x) {x[, colnames(x) %in% cnames]}
hn <- lapply(hn, keep_match)
rn <- lapply(rn, keep_match)
all_df <- rbind.data.frame(do.call(rbind, hn), do.call(rbind, rn))

# Set CO2 chamber conditions, match to actual Ca
clevels <- c(400, 300, 200, 100, 50, 20, 10, 
             500, 600, 700, 800, 900, 1000, 1200, 1500)
nearest_clevel <- Vectorize(function(x) {clevels[which.min(abs(x - clevels))]})

# set Qin chamber conditions, match to actual Qin
qlevels <- c(0, 50, 100, 200, 400, 
             600, 800, 1200, 1500)
nearest_qlevel <- Vectorize(function(x) {qlevels[which.min(abs(x - qlevels))]})

# Keep relevant columns
all <- all_df %>%
  select(Treatment, trt, ID, species, rep, obs, date, E, A, Ca, Ci, Pci, Pca, Qin, gsw) %>%
  mutate(roundCa = nearest_clevel(Ca),
         roundQin = nearest_qlevel(Qin))

all <- as.data.table(all)

# Functions to summarize
SE <- function(x) {sd(x, na.rm = T)/sqrt(sum(!is.na(x)))}
num <- function(x) {sum(!is.na(x))}

# Summarize A and gs for only when Ca = 400 and Qin = 1500, first by plant, then by treatment
all_sum <- all[roundCa == 400 & roundQin == 1500,
               .(A.m = mean(A,  na.rm = TRUE), A.s = SE(A), A.n = num(A),
                 gsw.m = mean(gsw, na.rm = TRUE), gsw.s = SE(gsw), gsw.n = num(gsw)),
               by = .(trt, ID, rep)]
all_sum2 <- all_sum[,.(A.m = mean(A.m,  na.rm = TRUE), A.s = SE(A.m), A.n = num(A.m),
                      gsw.m = mean(gsw.m, na.rm = TRUE), gsw.s = SE(gsw.m), gsw.n = num(gsw.m)),
                    by = .(trt)]

# Reformat from wide to narrow to wide again
ag_sum <- all_sum2 %>%
  pivot_longer(!trt, names_to = c("trait", "value."), names_pattern = "(.*).(.)") %>%
  pivot_wider(names_from = c("value."), values_from = c("value")) %>%
  rename(p.mean = m, p.se = s, p.n = n) %>%
  pivot_wider(names_from = trt, values_from = c(p.mean, p.se, p.n), names_sep = ".") 
  
```

Import parameters for all experiments and summarize. 

```{r}
in_params <- as.data.table(read.csv("~/sentinel-detection/data/derived_data/parameters_data.csv"))

params_sum <- in_params[Treatment %in% c("31_22_250", "31_31_250"), .(p.mean = mean(Value, na.rm = T), 
                            p.se = SE(Value),
                            p.n = num(Value)), by = .(Treatment, trait)]
p_sum <- params_sum %>%
  mutate(trt = case_when(Treatment == "31_22_250" ~ "Control",
                         Treatment == "31_31_250" ~ "High night T")) %>%
  select(-Treatment) %>%
  pivot_wider(names_from = trt, values_from = c(p.mean, p.se, p.n), names_sep = ".") %>%
  filter(trait != "g0M", trait != "g1M") %>%
  bind_rows(ag_sum) %>%
  mutate(across(2:5, round, 3))
```

## Create table

Format into pleasant output table using the 'gt' package. 

```{r}
p_gt <- p_sum[,1:5] %>%
  mutate(group = case_when(trait == "vmax" ~ "Photosynthetic",
                           trait == "LCPT" ~ "Photosynthetic",
                           trait == "LSP" ~ "Photosynthetic",
                           trait == "Am" ~ "Photosynthetic",
                           trait == "AQY" ~ "Photosynthetic",
                           trait == "Rd" ~ "Photosynthetic",
                           trait == "theta_lc" ~ "Photosynthetic",
                           trait == "A" ~ "Photosynthetic",
                           trait == "g0BB" ~ "Stomatal",
                           trait == "g1BB" ~ "Stomatal",
                           trait == "gsw" ~ "Stomatal",
                           )) %>%
  arrange(group, trait) %>%
  rename(Trait = "trait",
         mean.h = "p.mean.High night T",
         mean.c = "p.mean.Control",
         se.h = "p.se.High night T",
         se.c = "p.se.Control") %>%
  select(Trait, mean.c, se.c, mean.h, se.h)

tab1 <- gt(p_gt) %>%
  tab_spanner(
    label = md("**Control**"),
    columns = vars(mean.c, se.c)
  ) %>%
  tab_spanner(
    label = md("**High night T**"),
    columns = vars(mean.h, se.h)
  ) %>%
   tab_row_group(
    group = "Stomatal",
    rows = 9:11
  )  %>%
  tab_row_group(
    group = "Photosynthetic",
    row = 1:8
  ) %>%
  cols_label(
    mean.c = md("mean"),
    se.c = md("SE"),
    mean.h = md("mean"),
    se.h = md("SE")
  ) %>%
  tab_style(
    style = list(cell_text(style = "italic")
    ),
    locations = cells_row_groups()
  )
print(tab1)
gtsave(tab1, filename = "Table1.pdf", path = "/plots")
```

## Conduct Hotelling's multivariate t-test

Using A.m and gsw.m from all_sum and gas exchange parameters from in_params (SLA will be included when available for regular night temperature), format response matrix for Hotelling's multivariate t-test. 

```{r}

```

