---
title: "Phase II report figures"
author: "Jessica Guo"
date: "2/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Biomass and transpiration ts and diff figures

Create a panel of time-series and difference figures for the temperature comparison experiment. Use same colors as in 03_metaanalysis_figures.Rmd for the time series, but new colors for the differences. 
```{r}
library(readxl)
library(udunits2)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(cowplot)
```

First, prepare model output and validation data for plotting timeseries of biomass and transpiration. 
```{r}
# Organize 2 treatments into same figure
treatments <- c("rn", "hn")

biomass_ts <- c()
trans_ts <- c()
for (trt in treatments) {
  data_in <- read.csv(paste0("/data/output/pecan_runs/temp_comp_results/", trt, 
                        "/ensemble_ts_summary_AGB.csv")) %>%
    mutate(treatment = trt,
           trait = "biomass") %>%
    relocate(treatment)
  biomass_ts <- rbind(biomass_ts, data_in)
  
  data_in <- read.csv(paste0("/data/output/pecan_runs/temp_comp_results/", trt, 
                         "/ensemble_ts_summary_TVeg.csv")) %>%
    mutate(treatment = trt,
           trait = "transpiration") %>%
    relocate(treatment)
  trans_ts <- rbind(trans_ts, data_in)
}

# Bring in validation biomass
load("~/sentinel-detection/data/cleaned_data/biomass/chamber_biomass.Rdata")

biomass_valid <- chamber_biomass %>%
  filter(genotype == "ME034V-1" &
           treatment %in% c(NA, "control") &
           temp %in% c("31/22", "31/31") &
           light == 250 &
           !is.na(stem_DW_mg) &
           !is.na(leaf_DW_mg) &
           !is.na(panicle_DW_mg) &
           !is.na(root_DW_mg)) %>%
  rename(plant_id = plantID) %>%
  mutate(location = case_when(temp == "31/22" ~ "rn", 
                              temp == "31/31" ~ "hn"),
         stem_DW_g = ud.convert(stem_DW_mg, "mg", "g"),
         leaf_DW_g = ud.convert(leaf_DW_mg, "mg", "g"),
         panicle_DW_g = ud.convert(panicle_DW_mg, "mg", "g"),
         root_DW_g = ud.convert(root_DW_mg, "mg", "g")) %>%
  dplyr::select(location, plant_id, sowing_date, transplant_date, harvest_date,
         stem_DW_g, leaf_DW_g, panicle_DW_g, root_DW_g) %>%
  rename(treatment = location) %>%
  mutate(total_biomass_kg_m2 = ud.convert((stem_DW_g + leaf_DW_g + panicle_DW_g + root_DW_g)/103,
                                          "g/cm2", "kg/m2"),
         day = difftime(harvest_date, sowing_date, units = "days"),
         Trait = "Biomass~(kg~m^-2)",
         Treatment = case_when(treatment == "rn" ~ "Control night T",
                               treatment == "hn" ~ "High night T"))

# Combine both Biomass and Transpiration time series
all_ts <- rbind.data.frame(biomass_ts, trans_ts) %>%
  mutate(Trait = case_when(trait == "biomass" ~ "Biomass~(kg~m^-2)",
                           trait == "transpiration" ~ "Transpiration~(kg~m^-2~d^-1)"),
         Treatment = case_when(treatment == "rn" ~ "Control night T",
                               treatment == "hn" ~ "High night T"))

```

Plot first set of time series panels. 
```{r}
cols <- brewer.pal(7, name = "Set1")
fig_ts <- ggplot() +
  geom_line(data = all_ts, aes(day, y = median, color = Treatment)) +
  geom_ribbon(data = all_ts, aes(day, ymin = lcl_50, ymax = ucl_50, fill = Treatment), 
              alpha = 0.25) +
  # geom_point(data = biomass_valid, aes(x = day, y = total_biomass_kg_m2, color = Treatment)) +
  scale_x_continuous("Day of Experiment") + 
  facet_wrap(~Trait, ncol = 1, scales = "free_y", strip.position = "left", labeller = label_parsed) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        legend.title = element_blank(),
        legend.position = c(0.25, 0.9),
        legend.background = element_blank()) +
  scale_color_manual(name = NULL, values = cols[c(2,1)]) +
  scale_fill_manual(name = NULL, values = cols[c(2,1)])
```

Second, organize model output and plot differences between treatments.

```{r}
# Organize 2 treatments into same figure
biomass_df <- read.csv(paste0("/data/output/pecan_runs/temp_comp_results/comparison_diff_AGB.csv")) %>%
  pivot_wider(names_from = percentile, values_from = c(hn_rn)) %>%
  mutate(diff = "hn_rn",
         trait = "biomass") %>%
  rename(p025 = "25",
         p05 = "50",
         p25 = "250",
         p50 = "500",
         p75 = "750",
         p95 = "950",
         p975 = "975")

trans_df <- read.csv(paste0("/data/output/pecan_runs/temp_comp_results/comparison_diff_TVeg.csv")) %>%
  pivot_wider(names_from = percentile, values_from = c(hn_rn)) %>%
  mutate(diff = "hn_rn",
         trait = "transpiration") %>%
  rename(p025 = "25",
         p05 = "50",
         p25 = "250",
         p50 = "500",
         p75 = "750",
         p95 = "950",
         p975 = "975")

# Calculate when one-sided (.05) or two-sided (.025) t-test is significant
# Assumes that the difference is positive
biomass_df$sig_05 <- ifelse(biomass_df$p05 > 0, TRUE, FALSE)
biomass_df$sig_025 <- ifelse(biomass_df$p025 > 0, TRUE, FALSE)

trans_df$sig_05 <- ifelse(trans_df$p05 > 0, TRUE, FALSE)
trans_df$sig_025 <- ifelse(trans_df$p025 > 0, TRUE, FALSE)

all_df <- rbind.data.frame(biomass_df, trans_df) %>%
  mutate(Trait = case_when(trait == "biomass" ~ "Delta~Biomass~(kg~m^-2)",
                           trait == "transpiration" ~ "Delta~Transpiration~(kg~m^-2~d^-1)"),
         Diff = case_when(diff == "hn_rn" ~ "hn - cn"))

# Plot biomass treatment differences through time
col2 <- brewer.pal(3, name = "Dark2")
fig_diff <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(data = all_df, aes(day, y = p50, color = Diff)) +
  geom_ribbon(data = all_df, aes(day, ymin = p25, ymax = p75, fill = Diff), alpha = 0.25) +
  geom_point(data = all_df[all_df$sig_05 == TRUE,], aes(day, y = p50, color = Diff)) +
  scale_x_continuous("Day of Experiment") + 
  facet_wrap(~Trait, ncol = 1, scales = "free_y", strip.position = "left", labeller = label_parsed) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        legend.title = element_blank(),
        legend.position = c(0.2, 0.9),
        legend.background = element_blank()) +
  scale_color_manual(name = NULL, values = col2[3]) +
  scale_fill_manual(name = NULL, values = col2[3])
```

```{r}
jpeg(filename = "plots/Fig2_ts.jpg", height = 5, width = 7, units = "in", res = 600)
plot_grid(fig_ts, fig_diff, ncol = 2)
dev.off()
```

Finally, organize and plot variance decomposition for biomass and transpiration. 
```{r}
load("/data/output/pecan_runs/temp_comp_results/var_decomp_AGB.Rdata")
cols <- brewer.pal(7, name = "Set1")

# Exclude Leaf Respiration Rate and Leaf C:N
vd1 <- vd %>%
  filter(trait != "c2n_leaf" & trait != "growth_respiration_coefficient") %>%
  mutate_if(is.factor, as.character) %>%
  mutate(trait.labels = replace(trait.labels, trait == "extinction_coefficient_diffuse", 
                                  "Light Extinction Coefficient"),
         trait.labels = factor(trait.labels, levels = c("Vcmax", "Stomatal Slope BB", "Specific Leaf Area",
                                                        "Quantum Efficiency", "Light Extinction Coefficient",
                                                        "Leaf Respiration Rate", "Cuticular Conductance")),
         treatment = replace(treatment, treatment == "rn", "cn"))

cv.b <- ggplot(data = vd1) +
  geom_pointrange(aes(x = trait.labels, y = coef.vars, ymin = 0, ymax = coef.vars, 
                      col = treatment), 
                  alpha = 0.5, size = 0.75, position = position_dodge(width = c(-0.4))) +
  coord_flip() +
  geom_hline(aes(yintercept = 0), size = 0.1) +
  scale_y_continuous("CV (%)") +
  theme_classic(base_size = 10) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()) +
  scale_color_manual(name = NULL, values = cols[c(2,1)]) +
  guides(col = FALSE)

el.b <- ggplot(data = vd1) +
  geom_pointrange(aes(x = trait.labels, y = elasticities, ymin = 0, ymax = elasticities, 
                      col = treatment), 
                  alpha = 0.5, size = 0.75, position = position_dodge(width = c(-0.4))) +
  coord_flip() +
  scale_y_continuous("Elasticity") +
  geom_hline(aes(yintercept = 0), size = 0.1)  +
  theme_classic(base_size = 10) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank(), 
        axis.text.y=element_blank()) +
  scale_color_manual(name = NULL, values = cols[c(2,1)]) +
  guides(col = FALSE)

vdecomp.b <- ggplot(data = vd1) +
  geom_pointrange(aes(x = trait.labels, y = sd_convert, ymin = 0, ymax = sd_convert, 
                      col = treatment), 
                  alpha = 0.5, size = 0.75, position = position_dodge(width = c(-0.4))) +
  coord_flip() +
  geom_hline(aes(yintercept = 0), size = 0.1)  +
  scale_y_continuous(expression(paste("SD explained (kg ", m^-2, ")")), 
                     breaks = pretty(vd$sd_convert, n = 3)) +
  theme_classic(base_size = 10) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank(), 
        axis.text.y=element_blank(), 
        legend.position = c(0.9, 0.6),
        legend.title = element_blank(),
        legend.background = element_blank(),
        legend.key.size = unit(0.4, "cm")) +
  scale_color_manual(name = NULL, values = cols[c(2,1)])

# Plot variance decomposition of transpiration
load("/data/output/pecan_runs/temp_comp_results/var_decomp_TVeg.Rdata")

# Exclude Leaf Respiration Rate and Leaf C:N
vd2 <- vd %>%
  filter(trait != "c2n_leaf" & trait != "growth_respiration_coefficient") %>%
  mutate_if(is.factor, as.character) %>%
  mutate(trait.labels = replace(trait.labels, trait == "extinction_coefficient_diffuse", 
                                  "Light Extinction Coefficient"),
         trait.labels = factor(trait.labels, levels = c("Vcmax", "Stomatal Slope BB", "Specific Leaf Area",
                                                        "Quantum Efficiency", "Light Extinction Coefficient",
                                                        "Leaf Respiration Rate", "Cuticular Conductance")),
         treatment = replace(treatment, treatment == "rn", "cn"))
  
cv.t <- ggplot(data = vd2) +
  geom_pointrange(aes(x = trait.labels, y = coef.vars, ymin = 0, ymax = coef.vars, 
                      col = treatment), 
                  alpha = 0.5, size = 0.75, position = position_dodge(width = c(-0.4))) +
  coord_flip() +
  scale_y_continuous("CV (%)") +
  geom_hline(aes(yintercept = 0), size = 0.1) +
  theme_classic(base_size = 10) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()) +
  scale_color_manual(name = NULL, values = cols[c(2,1)]) +
  guides(col = FALSE)

el.t <- ggplot(data = vd2) +
  geom_pointrange(aes(x = trait.labels, y = elasticities, ymin = 0, ymax = elasticities, 
                      col = treatment), 
                  alpha = 0.5, size = 0.75, position = position_dodge(width = c(-0.4))) +
  coord_flip() +
  scale_y_continuous("Elasticity") +
  geom_hline(aes(yintercept = 0), size = 0.1)  +
  theme_classic(base_size = 10) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank(), 
        axis.text.y=element_blank()) +
  scale_color_manual(name = NULL, values = cols[c(2,1)]) +
  guides(col = FALSE)

vdecomp.t <- ggplot(data = vd2) +
  geom_pointrange(aes(x = trait.labels, y = sd_convert, ymin = 0, ymax = sd_convert, 
                      col = treatment), 
                  alpha = 0.5, size = 0.75, position = position_dodge(width = c(-0.4))) +
  coord_flip() +
  geom_hline(aes(yintercept = 0), size = 0.1)  +
  scale_y_continuous(expression(paste("SD explained (kg ", m^-2, " ", d^-1, ")")), 
                     breaks = pretty(vd$sd_convert, n = 3)) +
  theme_classic(base_size = 10) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank(), 
        axis.text.y=element_blank(), 
        legend.position = c(0.9, 0.6),
        legend.background = element_blank(),
        legend.key.size = unit(0.4, "cm")) +
  scale_color_manual(name = NULL, values = cols[c(2,1)])

jpeg(filename = "plots/Fig3_vardecomp.jpg", height = 5, width = 7, units = "in", res = 600)
cowplot::plot_grid(cv.b, el.b, vdecomp.b, cv.t, el.t, vdecomp.t, nrow = 2, rel_widths = c(1.5, 1, 1.1),
                   labels = c("Biomass", "", "", "Transpiration", "", ""),
                   align = "h", hjust = 0, vjust = 1,
                   label_size = 10)
dev.off()
```
