---
title: "Phase II report figures"
author: "Jessica Guo"
date: "2/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Biomass and transpiration ts and diff figures

Create a panel of time-series and difference figures for the 3 environments comparison. Use same colors as in 03_metaanalysis_figures.Rmd for the time series, but new colors for the differences. 
```{r}
library(readxl)
library(udunits2)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(cowplot)
```

First, prepare model output and validation data for plotting timeseries of biomass and transpiration. 
```{r}
# Organize 3 treatments into same figure
treatments <- c("ch", "gh", "out")

biomass_ts <- c()
trans_ts <- c()
for (trt in treatments) {
  data_in <- read.csv(paste0("/data/output/pecan_runs/env_comp_results/", trt, 
                        "/ensemble_ts_summary_TotLivBiom.csv")) %>%
    mutate(treatment = trt,
           trait = "biomass") %>%
    relocate(treatment)
  biomass_ts <- rbind(biomass_ts, data_in)
  
  data_in <- read.csv(paste0("/data/output/pecan_runs/env_comp_results/", trt, 
                         "/ensemble_ts_summary_TVeg.csv")) %>%
    mutate(treatment = trt,
           trait = "transpiration") %>%
    relocate(treatment)
  trans_ts <- rbind(trans_ts, data_in)
}

# Bring in validation biomass
load("~/sentinel-detection/data/cleaned_data/biomass/chamber_biomass.Rdata")
load("~/sentinel-detection/data/cleaned_data/biomass/greenhouse_outdoor_biomass.Rdata")

biomass_valid_ch <- chamber_biomass %>%
  filter(genotype == "ME034V-1" &
           treatment %in% c(NA, "control") &
           temp == "31/22" &
           light == 430 &
           !is.na(stem_DW_mg) &
           !is.na(leaf_DW_mg) &
           !is.na(panicle_DW_mg)) %>%
  rename(plant_id = plantID) %>%
  mutate(location = "ch",
         stem_DW_g = ud.convert(stem_DW_mg, "mg", "g"),
         leaf_DW_g = ud.convert(leaf_DW_mg, "mg", "g"),
         panicle_DW_g = ud.convert(panicle_DW_mg, "mg", "g")) %>%
  dplyr::select(location, plant_id, sowing_date, transplant_date, harvest_date,
         stem_DW_g, leaf_DW_g, panicle_DW_g)

biomass_valid_gh_out <- greenhouse_outdoor_biomass %>%
  filter(genotype == "ME034V-1" &
           treatment %in% c("pot", "jolly_pot") &
           !is.na(stem_DW_g) &
           !is.na(leaf_DW_g) &
           !is.na(panicle_DW_g) & 
           exp_site == "GH" & exp_number == 1 |
           genotype == "ME034V-1" &
           treatment %in% c("pot", "jolly_pot") &
           !is.na(stem_DW_g) &
           !is.na(leaf_DW_g) &
           !is.na(panicle_DW_g) & 
           exp_site == "Field" & exp_number == 2) %>%
  mutate(location = case_when(exp_site == "GH" ~ "gh",
                              exp_site == "Field" ~ "out")) %>%
  dplyr::select(location, plant_id, sowing_date, transplant_date, harvest_date,
         stem_DW_g, leaf_DW_g, panicle_DW_g)

# Combine 2 kinds of validation data
biomass_valid <- rbind(biomass_valid_ch, biomass_valid_gh_out) %>%
  rename(treatment = location) %>%
  mutate(total_biomass_kg_m2 = ud.convert((stem_DW_g + leaf_DW_g + panicle_DW_g)/103,
                                          "g/cm2", "kg/m2"),
         day = difftime(harvest_date, sowing_date, units = "days"))

# Combine both Biomass and Transpiration time series
all_ts <- rbind.data.frame(biomass_ts, trans_ts) %>%
  mutate(Trait = case_when(trait == "biomass" ~ "Biomass~(kg~m^-2)",
                           trait == "transpiration" ~ "Transpiration~(kg~m^-2~d^-1)"),
         Treatment = case_when(treatment == "ch" ~ "Chamber",
                               treatment == "gh" ~ "Greenhouse",
                               treatment == "out" ~ "Outdoor"))

```

Plot first set of time series panels. 
```{r}
cols <- brewer.pal(7, name = "Set1")
fig_ts <- ggplot() +
  geom_line(data = all_ts, aes(day, y = median, color = Treatment)) +
  geom_ribbon(data = all_ts, aes(day, ymin = lcl_95, ymax = ucl_95, fill = Treatment), 
              alpha = 0.25) +
  scale_x_continuous("Day of Experiment") + 
  facet_wrap(~Trait, ncol = 1, scales = "free_y", strip.position = "left", labeller = label_parsed) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        legend.title = element_blank(),
        legend.position = c(0.23, 0.9),
        legend.background = element_blank()) +
  scale_color_manual(name = NULL, values = cols[c(4,3,5)]) +
  scale_fill_manual(name = NULL, values = cols[c(4,3,5)])
```

Second, organize model output and plot differences between treatments.

```{r}
# Organize 3 treatments into same figure
biomass_df <- read.csv(paste0("/data/output/pecan_runs/env_comp_results/comparison_diff_TotLivBiom.csv")) %>%
  pivot_wider(names_from = percentile, values_from = c(gh_ch, out_ch, out_gh)) %>%
  pivot_longer(!day, names_to = c("first", "second", "level"), names_pattern = "(.*)_(.*)_(.*)",
               values_to = "percentile") %>%
  mutate(diff = paste0(first, "_", second),
         trait = "biomass") %>%
  pivot_wider(names_from = level, values_from = percentile) %>%
  rename(p025 = "25",
         p05 = "50",
         p50 = "500",
         p095 = "950",
         p975 = "975")

trans_df <- read.csv(paste0("/data/output/pecan_runs/env_comp_results/comparison_diff_TVeg.csv")) %>%
  pivot_wider(names_from = percentile, values_from = c(gh_ch, out_ch, out_gh)) %>%
  pivot_longer(!day, names_to = c("first", "second", "level"), names_pattern = "(.*)_(.*)_(.*)",
               values_to = "percentile") %>%
  mutate(diff = paste0(first, "_", second),
         trait = "transpiration") %>%
  pivot_wider(names_from = level, values_from = percentile) %>%
  rename(p025 = "25",
         p05 = "50",
         p50 = "500",
         p095 = "950",
         p975 = "975")

# Calculate when one-sided (.05) or two-sided (.025) t-test is significant
# Assumes that the difference is positive
biomass_df$sig_05 <- ifelse(biomass_df$p05 > 0, TRUE, FALSE)
biomass_df$sig_025 <- ifelse(biomass_df$p025 > 0, TRUE, FALSE)

trans_df$sig_05 <- ifelse(trans_df$p05 > 0, TRUE, FALSE)
trans_df$sig_025 <- ifelse(trans_df$p025 > 0, TRUE, FALSE)

all_df <- rbind.data.frame(biomass_df, trans_df) %>%
  mutate(Trait = case_when(trait == "biomass" ~ "Delta~Biomass~(kg~m^-2)",
                           trait == "transpiration" ~ "Delta~Transpiration~(kg~m^-2~d^-1)"),
         Diff = case_when(diff == "gh_ch" ~ "gh - ch",
                               diff == "out_ch" ~ "out - ch",
                               diff == "out_gh" ~ "out - gh"))

# Plot biomass treatment differences through time
col2 <- brewer.pal(3, name = "Dark2")
fig_diff <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(data = all_df, aes(day, y = p50, color = Diff)) +
  geom_ribbon(data = all_df, aes(day, ymin = p05, ymax = p095, fill = Diff), alpha = 0.25) +
  geom_point(data = all_df[all_df$sig_05 == TRUE,], aes(day, y = p50, color = Diff)) +
  scale_x_continuous("Day of Experiment") + 
  facet_wrap(~Trait, ncol = 1, scales = "free_y", strip.position = "left", labeller = label_parsed) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        legend.title = element_blank(),
        legend.position = c(0.2, 0.9),
        legend.background = element_blank()) +
  scale_color_manual(name = NULL, values = col2[c(1,3,2)]) +
  scale_fill_manual(name = NULL, values = col2[c(1,3,2)])
```

```{r}
jpeg(filename = "plots/Fig4_ts.jpg", height = 5, width = 7, units = "in", res = 600)
plot_grid(fig_ts, fig_diff, ncol = 2)
dev.off()
```

