---
title: "Fitting C4 Photosynthesis Model"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Information
## Edits by Tonantzin Tarin            Date/edits: Dec-2019

This script is taken from PEcAn repository 
It uses the Photosynthesis PEcAn package to fit A/Ci curvers
with Collatz et al. (1992) model for C4 plants. 

Here we are using LICO6800 data. Note that some variables have changed from LICOR6400 to LICOR6800, for example: photosynthesis was called "Photo" in LICOR6400, now is called "A" in the LICOR6400. 

Data is in the common LICOR format, but files has been cleaned up and two additional columns were added "species" and "rep". These stand for species (WT, ME034, A10) and replicates (plant_1, plant_2, ...n) respectively.

Raw data islocated in Cleaned_up_files folder 

```{r, message=FALSE, warning=FALSE}
library(PEcAn.photosynthesis)
library(rjags)
library(here)
library(readr)
library(dplyr)
library(tibble)
library(ggplot2)
knitr::opts_chunk$set(cache = TRUE)
```

## Load files

Here we are using one file at the time with three replicates (plant 1, plant 2 and plant 3).
This can be change to upload several files at the time.

Here we are also re-naming columns. 

```{r, message=FALSE, warning=FALSE}
new <- "A_Ci_curves_Sep052020_outdoor_JollyG"
dat <- read.csv(file.path(paste0("Cleaned_up_files/", new, ".csv")))

names(dat)[names(dat) == c("A")] <- "Photo"   # photosynthesis
names(dat)[names(dat) == c("Qin")] <- "PARi"  # light 
names(dat)[names(dat) == c("Pci")] <- "Ci_Pa" # intercellular CO2 concentration  
names(dat)[names(dat) == c("rep")] <- "id"    # plant ID
```

Reconfigure data into a list for each individual to run outlier detection

```{r, message=FALSE, warning=FALSE}
datlist <- list()
for(i in 1:length(unique(dat$id))){
  datlist[[i]] <- subset(dat, id==i)
}
```

Use the Licor_QC() function to remove outliers for each individual. 

```{r}
master <- list()
for(i in 1:length(datlist)){
    master[[i]] = Licor_QC(datlist[[i]], curve="ACi")
}
```

Combine into single dataframe and remove the flagged points

```{r}
dat <- do.call("rbind", master)
## if QC was done, remove both unchecked points and those that fail QC
if("QC" %in% colnames(dat)){
  dat <- dat[-which(dat$QC < 1),]  
} else {
  QC <- rep(1, nrow(dat))
  dat <- cbind(dat, QC)
}
```

Check data by plotting. 

```{r}
dat$id <- factor(dat$id)
fig_curves <- ggplot(dat)+
  geom_point(aes(x = Ci_Pa, y = Photo, col = id))+
  theme_bw(base_size = 12)
print(fig_curves)  
```

## Specify model

```{r}
my.model.jags <- "
model{
  alpha ~ dlnorm(-3.21,3.7) 	    	## initial slope of photosynthesis light response
  vmax ~ dlnorm(3,3)                ## maximum rubisco capacity
  r ~ dlnorm(-0.2,2.8)              ## leaf respiration
  k ~ dlnorm(11.5, 2.8)             ## initial slope of photosynthetic CO2 response
  tau ~ dgamma(0.1,0.1)

  for(i in 1:n){                ## process model
    al[i] <- alpha*q[i]         ## light limited without covariates
    ac[i] <- k*pi[i]/100000     ## CO2 limited without covariates
    ae[i] <- vmax               ## rubisco limited without covariates
    prean[i] <- min(min(al[i],ac[i]),ae[i])-r
    an[i] ~ dnorm(prean[i],tau) ## likelihood
    pA[i] ~ dnorm(prean[i],tau) ## prediction
  }
}
"
```

## Define initial conditions

```{r}
init <- list()
init[[1]] <- list(r = 0.8, vmax = 50,alpha = 0.03, tau = 10, k = 0.7*100000)
init[[2]] <- list(r = 1, vmax = 30, alpha = 0.07, tau = 20, k = 0.8*100000)
init[[3]] <- list(r = 2, vmax = 15,alpha = 0.06, tau = 20, k = 0.2*1000000)
```

## Fit the C4 model

Note here the number of iterations can be increased 

```{r, echo=TRUE, results='hide'}
spp.list = unique(as.character(dat$id))   
c4mcmc <- list()

for(s in spp.list){
  sel = which(dat$id == s)
  an = dat$Photo[sel]
  pi = dat$Ci_Pa[sel]
  q = dat$PARi[sel]

  c4 <- jags.model(file = textConnection(my.model.jags), 
                   data = list(an = an, pi = pi, q = q, n = length(an)), 
                   inits = init, 
                   n.chains = 3)
  update(c4, n.iter = 5000)
  c4.out <- coda.samples(model = c4, 
                         variable.names = c("r","vmax","alpha", "k", "prean", "pA"), 
                         n.iter = 100000, 
                         thin = 25)
  c4mcmc[[s]] = c4.out
}
```

## Review model fit

Here Rhat as a diagnostic factor for the fitting. It should be closer to 1. These will be saved in the "A_Ci_outputs" folder. 

```{r}
nom <- paste0("A_Ci_outputs/", new)
for(s in spp.list){
  print(gelman.diag(c4mcmc[[s]], multivariate = FALSE))
  pdf(paste(nom, s, "_gelman.pdf", sep = ""))
  gelman.plot(c4mcmc[[s]])
  dev.off()
}
```

## Save and plot output files

Density plots are save as PDFs in the "A_Ci_outputs" folder, and the comparison plots for predicted vs measured photosynthesis are printed. 

```{r}
nom <- paste0("A_Ci_outputs/", new)
for(s in spp.list){
  pdf(paste(nom, s, "_model_trace.pdf",sep=""))
  plot(as.mcmc.list(c4mcmc[[s]]))
  dev.off()
}


for(s in spp.list){
  sel1 = which(dat$id == s)
  an = dat$Photo[sel1]
  preans = data.frame(summary(c4mcmc[[s]])$statistics) %>% 
    select(Mean) %>% 
    rownames_to_column() %>% 
    filter(grepl("prean", rowname))
  
  pdf(paste(nom, s, "_FitvsObs.pdf", sep = ""))
  plot(an, preans$Mean, pch = 19, main = s, xlab = "Measured An (umol m-2 s-1)",
       ylab = "Predicted An (umol m-2 s-1)", cex.main = 1.3, cex.lab = 1.4)
  abline(0, 1, col="dark green", lwd = 3)
  dev.off()
}
```

## Results 

Select posterior statistics for vmax and save. 

```{r}
out <- data.frame()
for(i in 1:length(spp.list)){
  out <- rbind.data.frame(out, data.frame(summary(c4mcmc[[i]])$statistics) %>% rownames_to_column() %>% filter(rowname == "vmax"))
}
out$id <- spp.list

write.csv(out, paste0("A_Ci_outputs/", new, "_parameters.csv"), row.names = F)

```
